### 2020-12-16 experiments/eiv_max_likelihood.Rmd
First working commit. Univariate y(t) = A*(x(t+tau)) + C

---
title: "eiv_max_likelihood"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Define the problem.

```{r definition}
library("mvtnorm")
library("tidyverse")

a = 2
constant = 2
x = runif(100,0,10)
xi = x + rnorm(100,0,1)
plot(x, xi)
# Define a cubic function
y = a*x^2 + rnorm(100,0,10) + constant
plot(xi, y)

```

```{r}
# Plot the results of the regression without considering the EIV
lm(y~I(xi^2))
```
```{r}
# Define the solver functions
all_likelihoods = function(xi, y, x_theta, y_theta, alpha=alpha, c=c){

  shifts = NULL
  sum_l = 0
  for (i in 1:length(xi)){
    #print(i)
    point_l = point_likelihood(xi[i], y[i], x_theta, y_theta, alpha=alpha, c=c)
    #print(point_l)
    sum_l = sum_l + point_l$likelihood
    shifts = append(shifts, point_l$shift_z)
  }
  # Return the total shift likelihood across all points
  return(sum_l)
}

all_likelihoods_optim = function(params, xi, y){

  print(params)
  x_theta = params[1]
  y_theta=params[2]
  alpha=params[3]
  c=params[4]

  shifts = NULL
  sum_l = 0
  for (i in 1:length(xi)){
    #print(i)
    point_l = point_likelihood(xi[i], y[i], x_theta, y_theta, alpha=alpha, c=c)
    #print(point_l)
    sum_l = sum_l + point_l$likelihood
    shifts = append(shifts, point_l$shift_z)
  }

  print(sum_l)
  # Return the total shift likelihood across all points
  return(sum_l)
}

point_likelihood = function(xi0, yi, x_theta, y_theta, alpha=2, c=0){

  z_optim = optim(0, 
        fn=single_likelihood, 
        x_theta = x_theta,
        y_theta=y_theta, 
        yi=yi, 
        xi = xi0,
        alpha=alpha,
        c=c, 
        method="Brent",
        lower=-6*x_theta,
        upper=6*x_theta,
        type="mv")

  shift_z = z_optim$par
  max_l = z_optim$value

  return(list("shift_z"=(shift_z), "likelihood"=max_l))
}

single_likelihood = function(params, x_theta, y_theta, xi0, yi, alpha, c, type="individual"){

    z = params[1]
    xl = dnorm(z, 0, sd=x_theta, log=TRUE)
    y_p = alpha*(z + xi0)^2 + c
    res = yi - y_p
    yl = dnorm(res, 0, y_theta, log=TRUE)
    #print(paste0("XL: ",xl, " , YL: ", yl, ", Shift: ",z, ", Res: ",res))

    if (type == "individual"){
      return (-(yl + xl))
    } else {
      # Calculate the likelihood from a Multivariate Gaussian
      # Assume independence of the parameters
      covariance_mat = matrix(c(x_theta, 0, 0, y_theta), nrow=2, ncol=2)
      means = c(0,0)
      m_l = dmvnorm(c(z, res), means, covariance_mat, log = TRUE)
      return(-m_l)
    }
}
```

# Monte Carlo Simulation with Plots
```{r monte_carlo}
oy = c(1:100)/10
ox = c(1:100)/10
a = 1 + 1:20/10
c = c(1:100)/10

monte_carlo = as.data.frame(expand.grid(a, c, oy, ox))
monte_carlo$likelihood = NA
colnames(monte_carlo) = c("alpha","c","oy","ox","likelihood")
#monte_carlo$r = monte_carlo$ox/monte_carlo$oy

# Take a subset

monte_carlo = monte_carlo[sample(1:nrow(monte_carlo), 10000),]

for (i in 1:nrow(monte_carlo)){
  print(i)
  # Assume independent 
  l = all_likelihoods(xi, y, x_theta=monte_carlo[i,'ox'], y_theta=monte_carlo[i,'oy'], alpha=monte_carlo[i,'alpha'], c=monte_carlo[i,'c'])
  monte_carlo[i,'likelihood'] = l
}
```
```{r}
# Optimised version

model_fit = optim(c(1,1,1,1), fn=all_likelihoods_optim, xi=xi, y=y, method="L-BFGS")

```

```{r}
model_fit$par
```

```{r outcome}
monte_carlo = monte_carlo[order(monte_carlo$likelihood),]

plot(monte_carlo$ox, monte_carlo$likelihood, ylim=c(0,50))
```

```{r}
head(monte_carlo, 50)
plot(monte_carlo$ox, monte_carlo$likelihood, ylim=c(120, 1000))
```

```{r}
xt= (1:100)/10
y_p = 2.1*(xt)^2 - 15
plot(xt, y_p, ylim=c(-20, 250))
points(xi, y, col="red")
```

```{r}

```

### 2020-12-16 experiments/identifiability_of_eiv_systems.R

@@ -1,0 +1,291 @@
+library("mvtnorm")
+library("tidyverse")
+
+
+a = 2
+x = runif(10,0,10)
+xi = x + rnorm(10,0,2)
+plot(x, xi)
+y = a*x + rnorm(10,0,1)
+plot(xi, y)
+lm(y~xi)
+
+oy = c(1:100)/10
+ox = c(1:100)/10
+a = 1.5 + 1:10/10
+c = c(1:100)/10
+
+monte_carlo = as.data.frame(expand.grid(a, c, oy, ox))
+monte_carlo$likelihood = NA
+colnames(monte_carlo) = c("alpha","c","oy","ox","likelihood")
+monte_carlo$r = monte_carlo$ox/monte_carlo$oy
+
+# Known ratio of the standard deviations
+monte_carlo = subset(monte_carlo, monte_carlo$r == 2)
+
+# Remove a bunch to speed it up
+#monte_carlo = monte_carlo[sample(1:nrow(monte_carlo), size=50000),]
+
+
+points((1:100)/10, 1.6*((1:100)/10) + 3, col="red")
+points((1:100)/10, 1.3*((1:100)/10) + 4, col="blue")
+
+
+for (i in 1:nrow(monte_carlo)){
+  print(i)
+  # Assume independent 
+  l = all_likelihoods(xi, y, x_theta=monte_carlo[i,'ox'], y_theta=monte_carlo[i,'oy'], alpha=monte_carlo[i,'alpha'], c=monte_carlo[i,'c'])
+  monte_carlo[i,'likelihood'] = l
+}
+
+monte_carlo = monte_carlo[order(monte_carlo$likelihood),]
+
+plot(monte_carlo$ox, monte_carlo$likelihood, ylim=c(0,50))
+
+all_likelihoods = function(xi, y, x_theta, y_theta, alpha=alpha, c=c){
+  
+  shifts = NULL
+  sum_l = 0
+  for (i in 1:length(xi)){
+    #print(i)
+    point_l = point_likelihood(xi[i], y[i], x_theta, y_theta, alpha=alpha, c=c)
+    #print(point_l)
+    sum_l = sum_l + point_l$likelihood
+    shifts = append(shifts, point_l$shift_z)
+  }
+  # Return the total shift likelihood across all points
+  return(sum_l)
+}
+
+
+
[…]

### 2020-12-16 experiments/uniform_normal.R

@@ -1,0 +1,109 @@
+library("ggplot2")
+
+u = runif(1000,-20,20)
+n = rnorm(200,0,3)
+
+hist(c(u,n))
+
+
+
+b=(1:100)/10
+oy = (1:10)
+width=10:20
+width = width[width %% 2 == 0]
+ot=(1:30)/10
+ox=(1:2)
+#b=1
+#oy=1
+#width = 10
+#ot = 3
+#ox = 1
+mc = expand.grid(b, oy, width, ot, ox)
+mc$estimate_u = NA
+mc$estimate_sd = NA
+mc$t_u = NA
+mc$t_sd = NA
+colnames(mc)[1:5] = c("b","oy","width","ot","ox")
+estimate_sd = rep(NA, nrow(mc))
+t_sd = rep(NA, nrow(mc))
+# Shuffle
+mc = mc[sample(1:nrow(mc)),]
+for (j in 1:nrow(mc)){
+  print(j)
+  out = get_distributions(mc[j,'b'],
+                          mc[j,'oy'],
+                          mc[j,'width'],
+                          mc[j,'ot'],
+                          mc[j,'ox'])
+  
+  estimate_sd[j] = out$esd
+  t_sd[j] = out$tsd
+  #mc[j,'estimate_u'] = out$eu
+  #mc[j,'estimate_sd'] = out$esd
+  #mc[j,'t_u'] = out$tu
+  #mc[j,'t_sd'] = out$tsd
+  
+}
+
+plot(mc$ot/mc$oy, estimate_sd/t_sd)
+
+library(mgcv)
+mc$estimate_sd = estimate_sd
+mc$t_sd = t_sd
+gam_fit = mgcv::gam(estimate_sd/t_sd~s(b, k=3) + s(oy, k=3) + width + s(ot,k=3) + ox, data=mc)
+predict(gam_fit)
+
+
+get_distributions = function(b, oy, width, ot, ox){
+  # Select a random time point t between 1 & 20, by some shift distribution
+  estimated_shifts = NULL
+  ts = NULL
+  for (i in 1:1000){
+    y = rnorm(width + 1,0,oy)
+    t = round(rnorm(1,0,ot))
+    x = rnorm(1,0,ox)
+    y[width/2 + 1 + t] = y[width/2 + 1 + t] + b*x
+    res = y - b*x
+    
+    # Check for the likelihood
[…]

### 2020-12-16 linearObjective.py

@@ -1,0 +1,39 @@
+import src.support as src
+import numpy as np
+
+
+class linearTauSolver:
+    
+    
+    def __init__(self, X, y, A, tu, tsd, u, sd):
+        self.X = X
+        self.y = y
+        self.A = A
+        self.tu = tu
+        self.tsd = tsd
+        self.u = u
+        self.sd = sd
+    
+    
+    
+    def objective_function(self, shift_seq):
+        
+        X_shift = src.shift_array(self.X, shift_seq=np.array(shift_seq, dtype="int"))
+        X_shift = src.hor_mul(X_shift, self.A)
+        
+        # Create the prediction
+        y_p = np.sum(X_shift, axis=0)
+        
+        # Calculate the y axis residuals
+        res = self.y - y_p
+    
+        # Calculate the y axis log likelihood
+        e_l = src.log_likelihood(x=res, u=self.u, sd=self.sd)
+        # Calculate the t axis log likelihood
+        # The t Axis is a discrete parameter - using continuous likelihood leads to the parameter shrinkage
+        t_l = src.log_pmf_discrete(x=shift_seq, u=self.tu, sd=self.tsd)
+        
+        # Assume that the y and t errors are independent
+    
+        # Want to get the maximum likelihood! Hence the negative operator
+        return -(e_l + t_l)

### 2020-12-16 linearTVSM.py

@@ -1,0 +1,60 @@
+import src.support as src
+import numpy as np
+import scipy.optimize as o
+from linearObjective import linearTauSolver
+import rbfopt
+import os
+
+class linearTVSRModel:
+    
+    def __init__(self, rbf_settings, verbose=False):
+
+        
+        # Establish the settings
+        if rbf_settings is None:
+            self.settings = rbfopt.RbfoptSettings(max_evaluations=100, 
+                                 max_noisy_evaluations=10,
+                                 minlp_solver_path='/Users/aaronpickering/Desktop/bonmin-osx/bonmin',
+                                 print_solver_output=False)
+        
+        # Create a dev null output to silence the print output
+        if verbose is False:
+            self.dev_null = open(os.devnull, 'w')
+    
+
+    def inner_objective(self, params):
+                     
+        print(params)       
+        A = params[0]
+        tsd = params[1]
+        sd = params[2]
+        tu = 0
+        u = 0
+        f = linearTauSolver(self.X, self.y, A, tu, tsd, u, sd)
+        
+        # Create a user black box function
+        bb = rbfopt.RbfoptUserBlackBox(self.dimension, 
+                                       np.array([-2] * self.dimension), 
+                                       np.array([2] * self.dimension),
+                                       np.array(['I']*self.dimension), 
+                                       f.objective_function)
+
+        # Crreate the algorithm from the black box and settings
+        alg = rbfopt.RbfoptAlgorithm(self.settings,bb)
+        alg.set_output_stream(self.dev_null)
+        
+        val, x_out, itercount, evalcount, fast_evalcount = alg.optimize()
+        print("Likelihood: " + str(round(val,1)))
+        return val
+    
+    def fit(self, x_train, y_train):
+        self.x = x_train
+        self.X = src.decompose_vector(self.x)
+        self.dimension = self.X.shape[0]
+        self.y = y_train
+            
+        self.summary = o.minimize(self.inner_objective, [0, 2, 2], method='L-BFGS-B')
+        
+
+        
+        

### 2020-12-16 linearTVSR.py

@@ -1,0 +1,122 @@
+import src.support as src
+import numpy as np
+import scipy.optimize as o
+from linearObjective import linearTauSolver
+import rbfopt
+import os
+from scipy.optimize import dual_annealing
+from scipy.optimize import basinhopping
+from scipy.optimize import brute
+from scipy.optimize import fmin
+from scipy.optimize import differential_evolution
+import statsmodels.api as sm
+import pandas as pd
+import emcee
+from multiprocessing import Pool
+
+
+class linearTVSRModel:
+    
+    
+    
+    def __init__(self, rbf_settings, verbose=False):
+
+        self.shift_seq = None
+        self.likelihood = None
+        
+        self.posterior = pd.DataFrame({"A":0,"tsd":0,"sd":0,"likelihood":0}, index=[0])
+        
+        # Establish the settings
+        if rbf_settings is None:
+            self.settings = rbfopt.RbfoptSettings(max_evaluations=100, 
+                                 max_noisy_evaluations=10,
+                                 minlp_solver_path='/Users/aaronpickering/Desktop/bonmin-osx/bonmin',
+                                 print_solver_output=False)
+        else:
+            self.settings = rbf_settings
+        
+        # Create a dev null output to silence the print output
+        if verbose is False:
+            self.dev_null = open(os.devnull, 'w')
+            
+            
+    def basic_linear_regression(self):
+        # Statsmodels comparison (no intercept, same as above)
+        mod = sm.OLS(self.y, self.x)
+        res = mod.fit()
+        self.basic_lin_summary = res.summary()
+        self.min_A = res.params[0]
+        self.max_sd = np.std(res.predict(self.x)  - self.y)
+        
+    
+    def inner_objective(self, params):
+                     
+        print(params)       
+        A = params[0]
+        tsd = params[1]
+        sd = params[2]
+        
+        if tsd <= 0:
+            return 10000
+        
+        if sd <= 0:
+            return 10000
+        
+        tu = 0
+        u = 0
+        f = linearTauSolver(self.X, self.y, A, tu, tsd, u, sd)
+        
+        loop_bounds = src.refine_bounds(self.X_bounds.copy(), tsd)
+        print(loop_bounds)
+        
+        # Create a user black box function
+        bb = rbfopt.RbfoptUserBlackBox(self.dimension, 
+                                       loop_bounds[:,0], 
+                                       loop_bounds[:,1],
+                                       np.array(['I']*self.dimension), 
+                                       f.objective_function)
+
[…]

### 2020-12-16 main.py

@@ -1,0 +1,87 @@
+import pandas as pd
+import numpy as np
+import os
+import matplotlib.pyplot as plt
+import rbfopt
+from linearObjective import linearTauSolver
+import src.support as src
+from linearTVSR import linearTVSRModel
+import statsmodels.api as sm
+
+os.chdir("/Users/aaronpickering/Desktop/projects/cronus/")
+
+
+# Define additional variables
+A = np.array(8, dtype="float")
+np.random.seed(20)
+x = src.create_input(100, 0.2)
+X, bounds = src.decompose_vector(x, return_bounds=True)
+# Generate a random shift seq
+np.random.seed(21)
+real_shifts = np.round(np.random.randn(X.shape[0]))
+real_shifts = np.array(real_shifts, dtype="int")
+X_shift = src.shift_array(X, np.array(real_shifts, dtype="int"))
+X_shift = src.hor_mul(X_shift, A)
+
+xi = np.sum(src.shift_array(X, np.array(real_shifts, dtype="int")), axis=0)
+y = np.sum(X_shift, axis=0) + np.random.randn(X_shift.shape[1])
+
+tu=0
+tsd=1
+u=0
+sd=1
+
+# For comparison, the log likelihood of the actual maximum value (tau-error + y axis error)?
+f = linearTauSolver(X, y, 8, 0, 1, 0, 1)
+f.objective_function(real_shifts)
+
+
+
+# Plot the sequences
+plt.plot(y)
+plt.plot(x)
+plt.plot(xi)
+plt.show()
+
+# Define optimizer settings
+settings = rbfopt.RbfoptSettings(max_evaluations=100, 
+                                 max_noisy_evaluations=10,
+                                 minlp_solver_path='/Users/aaronpickering/Desktop/bonmin-osx/bonmin',
+                                 print_solver_output=False)
+
+
+## TVS Regression
+tvs = linearTVSRModel(settings)
+tvs.fit(x, y)
+# Print the summary
+tvs.summary
+tvs.basic_lin_summary
+tvs.shift_seq
+
+
+# Statsmodels comparison (no intercept, same as above)
+mod = sm.OLS(y, x)
+res = mod.fit()
+print(res.summary())
+res.params
+res.summary()
+
+
+# Plot the fit
+plt.scatter(x, y)
+plt.scatter(xi, y)
[…]

### 2020-12-16 src/support.py

@@ -1,0 +1,146 @@
+import pandas as pd
+import numpy as np
+from scipy.stats import norm
+from scipy.special import iv
+
+# Create the binomial input sequence
+def create_input(n, p = 0.1, binary=False):
+    ''' Creates a binomial vector of length n) '''
+    x = np.random.random(n)
+    x[x >= (1-p)] = 1
+    x[x < (1-p)]  = 0
+    
+    if binary is False:
+        x = np.random.randn(x.shape[0])*x
+    return x
+
+# Decompose a vector into the matrix representation
+def decompose_vector(x, return_bounds = False):
+    ''' Decompose a vector in a matrix representation, where
+    x = sum(i, n) xi 
+    The formula returns a matrix of the n vectors of xi '''
+    # Convert x to a numpy array
+    x = np.array(x)
+    x_out = np.zeros((x.shape[0], x.shape[0]))
+    bounds = np.zeros((x.shape[0], 2))
+
+    # Fill out the matrix
+    for i in range(0, x.shape[0]):
+        x_i = np.zeros((x.shape[0]))
+        x_i[i] = x[i]
+        x_out[i,:] = x_i
+        
+        # Fill the bounds array
+        bounds[i,0] = -1*i
+        bounds[i,1] = (x.shape[0] - i)
+
+    # Drop all the columns which are complete zeros
+    non_zero = np.apply_along_axis(sum , 0, x_out) != 0
+    x_out = x_out[non_zero,:]
+    bounds = bounds[non_zero, :]
+    if return_bounds is True:
+        return x_out, bounds
+    else:
+        return x_out
+
+
+
+
+# Define a lagging function
+def shift(x, shift=0):
+    x = np.roll(x, shift)
+    
+    if shift < 0:
+        x[shift:] = np.nan
+    elif shift > 0:
+        x[:shift ] = np.nan
+    
+    return x
+
+
+
+def shift_array(X, shift_seq, fill_na=True):
+    # Shift the values by the given sequence (this is actual shift value for inference)
+    X_shift = X.copy()
+    for i in range(0, X.shape[0]):
+        X_shift[i,:] = shift(X[i,:], shift=shift_seq[i])
+        
+    if fill_na is True:
+        X_shift[np.isnan(X_shift)] = 0
+        
+    return X_shift
+
+
+def hor_mul(X, A):
+    for i in range(0, X.shape[0]):
[…]

### 2020-12-16 linearTVSR.py
Added options for optimizer. Changed initial conditions to an easy example.

@@ -1,18 +1,14 @@
-import src.support as src
-import numpy as np
-import scipy.optimize as o
-from linearObjective import linearTauSolver
-import rbfopt
-import os
-from scipy.optimize import dual_annealing
-from scipy.optimize import basinhopping
-from scipy.optimize import brute
-from scipy.optimize import fmin
-from scipy.optimize import differential_evolution
-import statsmodels.api as sm
 import pandas as pd
-import emcee
-from multiprocessing import Pool
+import numpy as np
+import os
+import rbfopt
+from scipy.optimize import differential_evolution
+import scipy.optimize as o
+import statsmodels.api as sm
+
+# Local libraries
+import src.support as src
+from linearObjective import linearTauSolver
 
 
 class linearTVSRModel:
@@ -96,7 +92,7 @@ class linearTVSRModel:
         
         return val
     
-    def fit(self, x_train, y_train):
+    def fit(self, x_train, y_train, method="L-BFGS-B"):
         self.x = x_train
         self.X, self.X_bounds = src.decompose_vector(self.x, return_bounds=True)
         self.dimension = self.X.shape[0]
@@ -107,16 +103,17 @@ class linearTVSRModel:
         
         # Assign bounds
         bnds = ((self.min_A,10),(0.00,self.x.shape[0]/2), (0.00, self.max_sd))
-        #self.summary = o.minimize(self.inner_objective, 
-                                 # [10, 2, 2], 
-                                  #method='L-BFGS-B', bounds=bnds)
         
-
-        self.summary = differential_evolution(self.inner_objective,
-                                              bnds, 
-                                              polish=True,
-                                              maxiter=50,
-                                              popsize=1)
+        if method == "L-BFGS-B":
+            self.summary = o.minimize(self.inner_objective, [0, 2, 2], method='L-BFGS-B', bounds=bnds)
+        elif method == "differential_evolution":
+            self.summary = differential_evolution(self.inner_objective,
+                                                  bnds, 
+                                                  polish=True,
+                                                  maxiter=50,
+                                                  popsize=1)
+        else:
+            raise("Error: no compatible optimizer found")
 
         
         

### 2020-12-16 linearTVSM.py
Deleted unused file.

@@ -1,60 +1,0 @@
-import src.support as src
-import numpy as np
-import scipy.optimize as o
-from linearObjective import linearTauSolver
-import rbfopt
-import os
-
-class linearTVSRModel:
-    
-    def __init__(self, rbf_settings, verbose=False):
-
-        
-        # Establish the settings
-        if rbf_settings is None:
-            self.settings = rbfopt.RbfoptSettings(max_evaluations=100, 
-                                 max_noisy_evaluations=10,
-                                 minlp_solver_path='/Users/aaronpickering/Desktop/bonmin-osx/bonmin',
-                                 print_solver_output=False)
-        
-        # Create a dev null output to silence the print output
-        if verbose is False:
-            self.dev_null = open(os.devnull, 'w')
-    
-
-    def inner_objective(self, params):
-                     
-        print(params)       
-        A = params[0]
-        tsd = params[1]
-        sd = params[2]
-        tu = 0
-        u = 0
-        f = linearTauSolver(self.X, self.y, A, tu, tsd, u, sd)
-        
-        # Create a user black box function
-        bb = rbfopt.RbfoptUserBlackBox(self.dimension, 
-                                       np.array([-2] * self.dimension), 
-                                       np.array([2] * self.dimension),
-                                       np.array(['I']*self.dimension), 
-                                       f.objective_function)
-
-        # Crreate the algorithm from the black box and settings
-        alg = rbfopt.RbfoptAlgorithm(self.settings,bb)
-        alg.set_output_stream(self.dev_null)
-        
-        val, x_out, itercount, evalcount, fast_evalcount = alg.optimize()
-        print("Likelihood: " + str(round(val,1)))
-        return val
-    
-    def fit(self, x_train, y_train):
-        self.x = x_train
-        self.X = src.decompose_vector(self.x)
-        self.dimension = self.X.shape[0]
-        self.y = y_train
-            
-        self.summary = o.minimize(self.inner_objective, [0, 2, 2], method='L-BFGS-B')
-        
-
-        
-        

### 2020-12-16 main.py
Added -5, +5 bound on the shifts. This makes the optimisation smoother. In future, add a 'smart' boundary based on the ratio of the t-axis/y-axis likelihoods.

@@ -44,13 +44,15 @@ plt.show()
 
 # Define optimizer settings
 settings = rbfopt.RbfoptSettings(max_evaluations=100, 
-                                 max_noisy_evaluations=10,
+                                 max_noisy_evaluations=100,
                                  minlp_solver_path='/Users/aaronpickering/Desktop/bonmin-osx/bonmin',
                                  print_solver_output=False)
 
 
 ## TVS Regression
 tvs = linearTVSRModel(settings)
+
+# method is gradient descent like (L-BFGS-B) or genetic (differential_evolution)
 tvs.fit(x, y, method="L-BFGS-B")
 
 # Print the summary

### 2020-12-16 src/support.py

@@ -138,6 +138,12 @@ def refine_bounds(bounds, sd):
     t_min = np.min(bound_range)
     t_max = np.max(bound_range)
     
+    if t_min < -5:
+        t_min = -5
+        
+    if t_max > 5:
+        t_max = 5
+    
     # Refine the input bounds for each section
     bounds[bounds < t_min] = t_min
     bounds[bounds > t_max] = t_max

### 2020-12-17 main.py
Reorder folder structure

@@ -32,7 +32,7 @@ u=0
 sd=1
 
 # For comparison, the log likelihood of the actual maximum value (tau-error + y axis error)?
-f = linearTauSolver(X, y, 8, 0, 1, 0, 1)
+f = linearTauSolver(X, y, 2, 0, 1, 0, 1)
 f.objective_function(real_shifts)
 
 

### 2020-12-18 main.py
Split the inputs into semi-independent segments, so that the inner optimisation loop can be performed sequentially. This should speed up computation without a major loss of accuracy.

@@ -4,11 +4,11 @@ import os
 import matplotlib.pyplot as plt
 import rbfopt
 
-from linearObjective import linearTauSolver
-import src.support as src
-from linearTVSR import linearTVSRModel
 
-os.chdir("/Users/aaronpickering/Desktop/projects/cronus/")
+from src.model.linearObjective import linearTauSolver
+from src.model.linearTVSR import linearTVSRModel
+import src.support as src
+
 
 
 # Define additional variables

### 2020-12-18 src/model/linearObjective.py

@@ -13,6 +13,7 @@ class linearTauSolver:
         self.tsd = tsd
         self.u = u
         self.sd = sd
+
     
     
     

### 2020-12-18 src/model/linearTVSR.py

@@ -8,8 +8,8 @@ import statsmodels.api as sm
 
 # Local libraries
 import src.support as src
-from linearObjective import linearTauSolver
-
+from src.model.linearObjective import linearTauSolver
+from src.optimisation.dataSegments import dataSegments
 
 class linearTVSRModel:
     
@@ -60,43 +60,92 @@ class linearTVSRModel:
         
         tu = 0
         u = 0
-        f = linearTauSolver(self.X, self.y, A, tu, tsd, u, sd)
-        
+        # Refine bounds and define the objective function
         loop_bounds = src.refine_bounds(self.X_bounds.copy(), tsd)
-        print(loop_bounds)
         
-        # Create a user black box function
-        bb = rbfopt.RbfoptUserBlackBox(self.dimension, 
-                                       loop_bounds[:,0], 
-                                       loop_bounds[:,1],
-                                       np.array(['I']*self.dimension), 
-                                       f.objective_function)
 
-        # Crreate the algorithm from the black box and settings
-        alg = rbfopt.RbfoptAlgorithm(self.settings,bb)
-        alg.set_output_stream(self.dev_null)
+
+        ##### TO-DO
+        # Define an input to select whether to split or not
         
-        val, x_out, itercount, evalcount, fast_evalcount = alg.optimize()
+        segments = dataSegments(self.X.copy(), self.x.copy(), self.y.copy(), loop_bounds)
+        segments.refine_bounds()
         
+        # Loop through all the segments
+        vals = np.array([])
+        taus = np.array([])
+        print("gets here")
+        for j in range(0, len(segments.X_segments)):
+            X = segments.X_segments[j]
+            y = segments.y_segments[j]
+            dim = segments.dimension[j]
+            # Temporary fix until the bounds are better organised
+            bnds = 0
+ 
+            # Define our solver for the internal loop
+            f = linearTauSolver(X, y, A, tu, tsd, u, sd)
+            
+            # Run inner optimisation loop
+            val, tau = self.inner_optimisation(X, y, dim, bnds, f, loop_bounds)
+            print(val, tau)
+            vals = np.append(vals, val)
+            taus = np.append(taus, tau)
+        
+        max_likelihood = np.sum(vals)
         # Save the best state parameter from the internal loop
         if self.likelihood is None:
-            self.likelihood = val
-        elif val < self.likelihood:
-            self.likelihood = val
-            self.shift_seq = x_out
[…]

### 2020-12-18 src/optimisation/dataSegments.py

@@ -1,0 +1,84 @@
+import src.support as src
+import numpy as np
+
+
+class dataSegments(object):
+    
+
+    def __init__(self, X, x, y, bounds):        
+        self.loc, self.loc_diff = self.get_impulse_locations(x)
+        splits = self.get_split_points(bounds)
+        
+        # Split the dependent variable
+        self.y_segments = np.split(y, splits)
+        
+        # Split the input matrix
+        X_out = np.split(X, splits, axis=1)
+        # Remove elements with all zeroes
+        X_out = [z[z.any(axis=1)] for z in X_out]
+        # Assign the segments to the object
+        self.X_segments = X_out
+        
+        # Get the new updated object dimensions and bounds
+        self.dimension = [z.shape for z in X_out]
+
+        
+    
+    def get_impulse_locations(self, x):
+        
+        # Get the non zero element positions of the input vector
+        loc = np.where(x != 0)[-1]
+        loc_diff = np.diff(loc)
+        
+        return loc, loc_diff
+    
+    
+    def get_split_points(self, bounds):
+        # Extract the maximum boundary
+        self.global_max_bound = np.max(bounds)
+        self.global_min_bound = np.min(bounds)
+        # Find the spaces which are greater than twice the boundary (2 points movin toward each other)
+        splits = np.where(self.loc_diff > np.max([np.abs(self.global_min_bound), self.global_max_bound])*2)
+        # Fit the array location where these splits occur
+        split_locations = self.loc[splits] + np.round(self.loc_diff[splits]/2)
+        
+        return split_locations.astype("int16")
+    
+    
+    def refine_bounds(self):
+        
+        # Get the cumulative sum of the various split locations
+        cum_sum = [0]
+        for i in range(0, len(self.dimension)):
+                cum_sum.append(self.dimension[i][1] + cum_sum[i])
+        cum_sum = np.array(cum_sum)
+        
+        # Loop through the absolute values and refine the global boundaries
+        #   into local values that change per split
+        
+        self.bounds = []
+        for l in self.loc:
+            diff = l - cum_sum
+            min_bound = -1*np.min(diff[diff > 0])
+            max_bound = -1*np.max(diff[diff < 0])
+            # Check against the global max bound
+            if max_bound > self.global_max_bound:
+                max_bound = self.global_max_bound
+            if min_bound < self.global_min_bound:
+                min_bound = self.global_min_bound 
+            # Add to the output array
+            np.append(self.bounds, [min_bound, max_bound])
[…]

### 2020-12-18 src/optimisation/segmentation.py

@@ -1,0 +1,85 @@
+import src.support as src
+import numpy as np
+
+
+test = dataSegments(X, x, bounds)
+test.refine_bounds()
+
+class dataSegments(object):
+    
+
+    def __init__(self, X, x, bounds):        
+        self.loc, self.loc_diff = self.get_impulse_locations(x)
+        splits = self.get_split_points(bounds)
+        
+        X_out = np.split(X, splits, axis=1)
+        
+        # Remove elements with all zeroes
+        X_out = [z[z.any(axis=1)] for z in X_out]
+        
+        # Assign the segments to the object
+        self.segments = X_out
+        # Get the new updated object dimensions and bounds
+        self.dimension = [z.shape for z in X_out]
+
+        
+    
+    def get_impulse_locations(self, x):
+        
+        # Get the non zero element positions of the input vector
+        loc = np.where(x != 0)[-1]
+        loc_diff = np.diff(loc)
+        
+        return loc, loc_diff
+    
+    
+    def get_split_points(self, bounds):
+        # Extract the maximum boundary
+        self.global_max_bound = np.max(bounds)
+        self.global_min_bound = np.min(bounds)
+        # Find the spaces which are greater than twice the boundary (2 points movin toward each other)
+        splits = np.where(self.loc_diff > np.max(np.abs(self.global_min_bound), self.global_max_bound)*2)
+        # Fit the array location where these splits occur
+        split_locations = self.loc[splits] + np.round(self.loc_diff[splits]/2)
+        
+        return split_locations.astype("int16")
+    
+    
+    def refine_bounds(self):
+        
+        # Get the cumulative sum of the various split locations
+        cum_sum = [0]
+        for i in range(0, len(self.dimension)):
+                cum_sum.append(self.dimension[i][1] + cum_sum[i])
+        cum_sum = np.array(cum_sum)
+        
+        # Loop through the absolute values and refine the global boundaries
+        #   into local values that change per split
+        
+        self.bounds = []
+        for l in self.loc:
+            diff = l - cum_sum
+            print(diff)
+            min_bound = -1*np.min(diff[diff > 0])
+            max_bound = -1*np.max(diff[diff < 0])
+            # Check against the global max bound
+            if max_bound > self.global_max_bound:
+                max_bound = self.global_max_bound
+            if min_bound < self.global_min_bound:
+                min_bound = self.global_min_bound 
+            # Add to the output array
+            self.bounds.append((min_bound, max_bound))
+                
+                
[…]

### 2020-12-18 main.py
Iterators now work correctly. Need to treat edge case, tsd~=0

@@ -43,8 +43,8 @@ plt.plot(xi)
 plt.show()
 
 # Define optimizer settings
-settings = rbfopt.RbfoptSettings(max_evaluations=100, 
-                                 max_noisy_evaluations=100,
+settings = rbfopt.RbfoptSettings(max_evaluations=10, 
+                                 max_noisy_evaluations=10,
                                  minlp_solver_path='/Users/aaronpickering/Desktop/bonmin-osx/bonmin',
                                  print_solver_output=False)
 

### 2020-12-18 src/model/linearObjective.py

@@ -19,15 +19,18 @@ class linearTauSolver:
     
     def objective_function(self, shift_seq):
         
+        
         X_shift = src.shift_array(self.X, shift_seq=np.array(shift_seq, dtype="int"))
+
         X_shift = src.hor_mul(X_shift, self.A)
+        
         
         # Create the prediction
         y_p = np.sum(X_shift, axis=0)
         
         # Calculate the y axis residuals
         res = self.y - y_p
-    
+
         # Calculate the y axis log likelihood
         e_l = src.log_likelihood(x=res, u=self.u, sd=self.sd)
         # Calculate the t axis log likelihood

### 2020-12-18 src/model/linearTVSR.py

@@ -74,8 +74,9 @@ class linearTVSRModel:
         # Loop through all the segments
         vals = np.array([])
         taus = np.array([])
-        print("gets here")
+
         for j in range(0, len(segments.X_segments)):
+            print("Segment " + str(j) + " from " + (str(len(segments.X_segments) - 1)))
             X = segments.X_segments[j]
             y = segments.y_segments[j]
             dim = segments.dimension[j]
@@ -87,7 +88,7 @@ class linearTVSRModel:
             
             # Run inner optimisation loop
             val, tau = self.inner_optimisation(X, y, dim, bnds, f, loop_bounds)
-            print(val, tau)
+
             vals = np.append(vals, val)
             taus = np.append(taus, tau)
         
@@ -124,8 +125,8 @@ class linearTVSRModel:
             val = f.objective_function(x_out)
         else:
             bb = rbfopt.RbfoptUserBlackBox(dim[0], 
-                                           [-4]*dim[0], 
-                                            [4]*dim[0],
+                                           np.array([-4]*dim[0]), 
+                                            np.array([4]*dim[0]),
                                             np.array(['I']*dim[0]), 
                                             f.objective_function)
     

### 2021-02-02 documentation/article/article.tex
add initial version of introduction

\documentclass[11pt]{amsart}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[all]{xy}
\usepackage{graphicx}

\begin{document}

\title{Modelling Stochastic Time Delay Structures for Regression}
\author{Aaron Pickering}
%\address{}
\email{aaron1rcl@gmail.com}
%\urladdr{} 
\date{\today}
\maketitle

\section{Introduction}

Consider the typical univariate regression problem where the intention is to find
the relationship between $x$ and $y$. When extended to time series, a number of time specific complexities arise. For the specific case where the input $x$ affects the output $y$, with a random time delay in between, the estimated coefficients (or weights, predictions etc) are significantly attenuated. This attenuation occurs both in traditional statistical models and machine learning models, and for certain problems can be a serious limitation. The following document proposes techniques for handling this class of time series regression.

As an example, let us take the management of blood sugar level in diabetes patients. In people with diabetes, the pancreas can't effectively regulate blood sugar levels. Therefore, these levels must be controlled by insulin injections and a special diet. The challenge for many people is that the relationship between the input (insulin) and output (blood sugar) is extremely complex. The effect of the insulin injection might be observed after $15$, $20$  or $t$ minutes depending on a number of factors, many of which are unknown. Due to the stochastic nature of the time delays, the actual effect can't be easily determined. It is difficult to differentiate the effect of the insulin injection from other factors and accurately determine how much one should take.

Inference for this type of problem is especially challenging. Typical regression models require a fixed alignment between cause and effect. Using standard methods, we'd need to assume that the effect occurs after some fixed time 't' which can be inferred from the data. However, if there is any uncertainty in the parameter $t$ ($t$ changes or is noisy) the resulting estimates will be significantly attenuated.

Consider the simple example where the effect of the input is $1$, as shown in the image below. The observed input is given by the red line, the blue line is when the effect actually occurs. The first effect happens 1 time point after the input. The second effect happens at the same time as the input. A fixed time delay isn't valid in this case because the time shifts differ.

\begin{center}
\begin{figure}
\includegraphics[totalheight=8cm]{images/1_input_delay_example.png}
\caption{Add figure caption}
\label{fig:verticalcell}
\end{figure}
\end{center}

If one were to model the effect using a fixed time delay and OLS, the estimate would only be half the true value because only one of the outputs is aligned. Obviously, this isn't ideal, we want the parameter estimates to be as close to the real values as possible, regardless of any noise in the lag structures. One can mitigate the problem via time aggregations, however in complicated multivariate cases, this is just not feasible.

Therefore, we propose a regression model which can handle stochastic time delay structures. We treat the stochastic time delay components as an ‘error’ in the time axis. Next, we find the maximum likelihood estimate, for a given set of parameters, considering the time error ($t$-axis) and the regression error ($y$-axis) simultaneously.

\section{Methodology}

...

\end{document}

### 2021-02-03 documentation/article/article.tex
add fist version methodology

@@
 \usepackage{graphicx}
+\usepackage{xcolor}
 
@@
 
-Consider the typical univariate regression problem where the intention is to find
-the relationship between $x$ and $y$.
+Consider the typical univariate regression problem where the intention is to find the relationship between $x$ and $y$.
 When extended to time series, a number of time specific complexities arise.
@@
 Typical regression models require a fixed alignment between cause and effect.
-Using standard methods, we'd need to assume that the effect occurs after some fixed time 't' which can be inferred from the data.
+Using standard methods, we'd need to assume that the effect occurs after some fixed time $t$ which can be inferred from the data.
 However, if there is any uncertainty in the parameter $t$ ($t$ changes or is noisy) the resulting estimates will be significantly attenuated.
@@
 The observed input is given by the red line, the blue line is when the effect actually occurs.
-The first effect happens 1 time point after the input.
+The first effect happens one time point after the input.
 The second effect happens at the same time as the input.
@@
 \begin{figure}
-\includegraphics[totalheight=8cm]{images/1_input_delay_example.png}
+\includegraphics[totalheight=6cm]{images/2_input_delay_example.png}
 \caption{Add figure caption}
@@
 
-...
+We consider the problem as analogous to the typical error-in-variables (EIV) regression.
+Ordinary regression analyses (and machine learning models) define the loss function with respect to errors in the y axis only.
+For EIV, errors are considered in both the $y$-axis and the $x$-axis (REF).
+This is useful when there are measurement errors in the independent variable eg.
+because your physical measurements have some degree of random error.
+Similarly, for this problem we assume that we have errors in the $y$-axis and the $t$-axis.
+That is, there are random prediction errors and random errors in the time domain.
+
+Let $x: [a, b] \longrightarrow \mathbb{R}$ be a discrete time series with finite domain and we want to determine the injective, functional relationship $f$ to some other variable y.
+The series $x(t)$ must be a stationary series with a known value for the point at which $f(x(t)) = 0$.
+That is, we know at which point the input series has no effect on the output (when it’s off).
+
+% TODO: Rewrite concrete expression of support...is it on x ot t?
+\textcolor{red}{For clarity, as per convention, we use the nomenclature ‘support’ for the elements $f(x(t)) \neq 0$}.
+In addition, the series $x(t)$ should be sparse, in that it contains a low density of ‘non-zero’ coordinates.
+
+We also assume that the set of $\tau$'s, denoted $\mathcal{T}$, is not a constant, but rather a random draw from some discrete distribution (e.g.
+discrete gaussian, poisson etc).
+For the discrete gaussian kernel $\mathcal{T}\sim T(0, \sigma)$ or for the poisson distribution $\mathcal{T}\sim Pois(\lambda)$ and $\varepsilon \sim N(0, \sigma_{\varepsilon})$.
+
+Firstly, let us take the input series $x(t)$ and decompose it into its constituent non-zero impulse components. So, if $x(t)$ is a vector given by
+
+\begin{align}
+x(t) = \left(0,0,1,0,1,0\right)
+\end{align}
+
+then we decompose the vector into a matrix
+
+\begin{align}
+x(t) = 
+\left(
+\begin{array}{cccccc}
+0 & 0 & 1 & 0 & 0 & 0 \\
+0 & 0 & 0 & 0 & 1 & 0 
+\end{array}  
+\right)
+\end{align}
+
+where each impulse is treated separately. \\
+
+Our algorithm then attempts to find the maximum likelihood estimate for the following equation:
+
+% TODO fix notation for T
+\begin{align}
+y(t)= f(X(t + T)) + \varepsilon
+\end{align}
+
+where $X(t + T)$ is the time shifted series and $X(t)$ is the observed input. \\
+
+In order to find the best estimate of f, we would like to find the function $f$ which maximises the joint likelihood of both Tau and e, the time-domain likelihood and the error likelihood respectively.
+
+Specifically, we maximise the log-likelihood of the following relation:
+% TODO: Clean this notation.
+\begin{align}
+\max  \left( \log \left(\mathcal{L}(\theta_{\varepsilon}, \varepsilon) \times \mathcal{L}(\theta_{\tau}, \tau)\right) \right)
+\end{align}
+
+where $\theta_{\varepsilon}$ and $\theta_{\tau}$ represent the parameters of the prediction error and time error respectively.
+For simplicity, assume that the tau and error distributions are independent.
+To begin with, the values of each individual time shift ($\tau$) are not known.
+In addition, the size of prediction error $\varepsilon$ can only be determined if each tau is known (because for each time shift there is a different prediction and prediction error).
+Therefore, our algorithm finds the best estimate of the individual time shifts in an inner optimisation, then iteratively searches for the best estimate of f in an outer optimisation step.
+\\
 

### 2021-02-04 src/model/linearTVSR.py
Poisson model integration

@@ -34,6 +34,7 @@ class linearTVSRModel:
     shift_seq = None
     likelihood = None
     history = None
+    family = "gaussian"
 
     
     def __init__(self, verbose=False):
@@ -157,7 +158,7 @@ class linearTVSRModel:
                 dim_seg = dim_segs[j]
 
                 # Define our solver for the internal loop
-                f = linearTauSolver(X_seg, y_seg, A, tu, tsd, u, sd, C)
+                f = linearTauSolver(X_seg, y_seg, A, tu, tsd, u, sd, C, self.family)
                 
                 # Run inner optimisation loop
                 val, tau = self.inner_optimisation(dim_seg, f)
@@ -167,7 +168,7 @@ class linearTVSRModel:
                 taus = np.append(taus, tau)
         else:
             # Define the solver function
-            input_obj = linearTauSolver(np.copy(self.X), np.copy(self.y), A, tu, tsd, u, sd, C)
+            input_obj = linearTauSolver(np.copy(self.X), np.copy(self.y), A, tu, tsd, u, sd, C, self.family)
             # Run the inner optimisation.
             vals, taus = self.inner_optimisation(np.copy(self.X.shape), input_obj)
 
@@ -214,19 +215,26 @@ class linearTVSRModel:
             
     def inner_optimisation(self,  dim, f):
         # Run the inner optimisation algorithm and retrieve the max likelihood
-        val, tau = tau_optimiser([0]*dim[0], f, 1000, 3)
+        val, tau = tau_optimiser([0]*dim[0], f, 1000, 3, self.family)
 
         return val, tau
             
             
-    def fit(self, x_train, y_train, method="L-BFGS-B", split=True, max_t=None):
+    def fit(self, x_train, y_train, method="L-BFGS-B", split=True, max_t=None, family=None):
         
         # Define the setting
         self.split = split
         
+        # Set the family
+        if family is not None:
+            self.family = family
+        
         # Load the data into the object memory and decompose the vector
         self.x = x_train
         self.y = y_train
+        
+        # Run a standard linear regression first for comparison
+        self.basic_linear_regression()
         
         # Scale the input parameter space and save the transforms (makes optimisation easier)
         self.scale_inputs(max_t=max_t)
@@ -236,8 +244,7 @@ class linearTVSRModel:
         self.dimension = self.X.shape[0]
 
         
-        # Run a standard linear regression first
-        self.basic_linear_regression()
+
         
         # Assign bounds
         # Bounds can be fixed because the input space has been scaled

### 2021-02-05 src/distributions.py
Added support for Poisson distribution time errors (#19)
* Linear objective for poisson time error.

* Poisson model integration

* Poisson integration in optimizer.

* Poisson distribution additions.

* Moved distributions into separate file.

@@ -1,0 +1,20 @@
+import numpy as np
+from scipy.stats import poisson
+from scipy.special import iv
+
+
+
+# Modified bessel function from wikipedia for discrete gaussian case
+def discrete_gaussian_kernel(x, sd):
+    try:
+        return np.exp(-sd) * iv(x, sd)
+    except:
+        return 1000
+    
+def discrete_poisson_kernel(x, mu):
+    try:
+        dist = poisson(mu)
+        return np.array(dist.pmf(x))
+    except:
+        return 1000
+    

### 2021-02-05 src/model/linearObjective.py

@@ -5,7 +5,7 @@ import numpy as np
 class linearTauSolver:
     
     
-    def __init__(self, X, y, A, tu, tsd, u, sd, C):
+    def __init__(self, X, y, A, tu, tsd, u, sd, C, family):
         self.X = np.copy(X)
         self.y = np.copy(y)
         self.A = A
@@ -14,8 +14,8 @@ class linearTauSolver:
         self.u = u
         self.sd = sd
         self.C = C
+        self.family = family
         
-    
     
     
     def objective_function(self, shift_seq):
@@ -45,11 +45,14 @@ class linearTauSolver:
         e_l = src.log_likelihood(x=res, u=self.u, sd=self.sd)
         # Calculate the t axis log likelihood
         # The t Axis is a discrete parameter - using continuous likelihood leads to the parameter shrinkage
-        t_l = src.log_pmf_discrete(x=shift_seq, u=self.tu, sd=self.tsd)
+        t_l = src.log_pmf_discrete(x=shift_seq, u=self.tu, sd=self.tsd, family=self.family)
         
 
         # Want to get the maximum likelihood! Hence the negative multiplier
         return -(e_l + t_l)
+    
+
+
     
     
     def predict(self, shift_seq):

### 2021-02-05 src/optimisation/custom_tau_optimiser.py

@@ -1,9 +1,7 @@
 import numpy as np
-# Define the optimizer
 
 
-
-def tau_optimiser(init, f, max_iterations, max_dependency=4):
+def tau_optimiser(init, f, max_iterations, max_dependency, family):
   
     '''
     Optimizer for best Tau Selection
@@ -58,7 +56,12 @@ def tau_optimiser(init, f, max_iterations, max_dependency=4):
         # The random tau selection controlled by the tau sd given on line 30
         proposal = np.copy(init)
         for sel in selection:
-            flip = np.round(np.random.randn(1)*f.tsd)
+            if family == "gaussian":
+                flip = np.round(np.random.randn(1)*f.tsd)
+            elif family == "poisson":
+                flip = np.random.poisson(f.tsd)
+            else:
+                raise("Error: family not found")
             proposal[sel] = flip
         # Check whether the proposal is better than the initial
         

### 2021-02-05 src/support.py

@@ -1,8 +1,11 @@
 import numpy as np
 from scipy.stats import norm
 from scipy.stats import mode
-from scipy.special import iv
 import itertools
+
+# Import custom distributions
+from src.distributions import discrete_gaussian_kernel
+from src.distributions import discrete_poisson_kernel
 
 # Create the binomial input sequence
 def create_input(n, p = 0.1, binary=False):
@@ -80,7 +83,15 @@ def hor_mul(X, A):
         X[i,:] =  X[i,:] * A
     return X
 
+
 def log_likelihood(x, u, sd):
+    ''' Calculate continous PMF log likelihood 
+        Currently setup only for gaussian error
+        Inputs:
+            x: the individual observations
+            u: gaussian mean
+            sd: gaussian standard deviation
+    '''
     l = np.log(norm.pdf(x,loc=u, scale=sd))
     
     if np.all(np.isnan(l)):
@@ -90,12 +101,24 @@ def log_likelihood(x, u, sd):
 
     return np.sum(l)
 
-def log_pmf_discrete(x, u, sd):
+def log_pmf_discrete(x, u, sd, family="gaussian"):
+    ''' Calculate the discrete PMF log likelihood for given data points
+        Inputs:
+            x: individual observations
+            u: ** currently set to zero - need to generalise **
+            sd: standard deviation for gaussian, mu for poisson
+            family: either gaussian or poisson
+    '''
     x = np.array(x)
     # Centre the variable first by subtracting the mean (to make it mean zero)
     x = x - u
     # Get the log likelihood
-    l = np.log(discrete_gaussian_kernel(x, sd))
+    if family == "gaussian":
+        l = np.log(discrete_gaussian_kernel(x, sd))
+    elif family == "poisson":
+        l = np.log(discrete_poisson_kernel(x, sd))
+    else:
+        raise("Error: family not found")
     
     if np.all(np.isnan(l)):
         return -100
@@ -104,31 +127,10 @@ def log_pmf_discrete(x, u, sd):
     l = l[~np.isnan(l)]
     return np.sum(l)
 
-def linear_objective_fn(X, y, shift_seq, A, tu, tsd, u, sd):
-    
-    X = shift_array(X, shift_seq=shift_seq)
-    
-    # Multiply the array vectors by the A
-    for i in range(0, X.shape[0]):
-        X[i,:] =  X[i,:] * A
-    
[…]

### 2021-02-06 documentation/article/article.tex
added v1 algorithm

@@
 \begin{align}
-x(t) = \left(0,0,1,0,1,0\right)
+x(t) = 
+\left(
+\begin{array}{cccccc}
+0 & 0 & 1 & 0 & 1 & 0
+\end{array}  
+\right)
 \end{align}
@@
 
-where $X(t + T)$ is the time shifted series and $X(t)$ is the observed input. \\
+where $X(t + T)$ is the time shifted series and $X(t)$ is the observed input.\\
 
-In order to find the best estimate of f, we would like to find the function $f$ which maximises the joint likelihood of both Tau and e, the time-domain likelihood and the error likelihood respectively.
+In order to find the best estimate of $f$, we would like to find the function $f$ which maximises the joint likelihood of both Tau(?) and $\varepsilon$, the time-domain likelihood and the error likelihood respectively.
 
@@
 In addition, the size of prediction error $\varepsilon$ can only be determined if each tau is known (because for each time shift there is a different prediction and prediction error).
-Therefore, our algorithm finds the best estimate of the individual time shifts in an inner optimisation, then iteratively searches for the best estimate of f in an outer optimisation step.
-\\
+Therefore, our algorithm finds the best estimate of the individual time shifts in an inner optimisation, then iteratively searches for the best estimate of $f$ in an outer optimisation step.
+
+\section{Algorithm}
+
+Firstly, we define a parametric form and some initial parameters to be estimated for the function $f$.
+For example, let's take the simple univariate linear model where the $f(x(t))$ is parameterised by $\beta$, $\sigma_{\varepsilon}$(error standard deviation)  and $\sigma_{\tau}$ (time error standard deviation).
+First, we initialise some starting values for each of these parameters.
+
+Now we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$.
+It stands to reason that the best possible time shift would be one that is not too far away from the observed impulse and also gives the best possible prediction.
+In this example, we can get the prediction y by simply multiplying the shifted value by its parameter $\beta$.
+From there we can calculate the likelihood estimate for time shift plus the prediction error.
+
+In principle, we can then try a number of values of $\tau$ (i.e. optimise) until we maximise the likelihood for this impulse.
+
+However, we must also consider that the impulses in $X(t)$ are not independent from each other. After shifting, it's possible that two or more effects can occur simultaneously.
+
+This could be particularly problematic if there are multiple impulses within a short period of time, or the impulses have a distributed effect over multiple time points.
+
+As an example consider the series
+
+\begin{align}
+x(t) = 
+\left(
+\begin{array}{cccccc}
+0 & 0 & 1 & 1 & 0 & 0
+\end{array}  
+\right)
+\end{align}
+
+with $T= (1, 0)$ and $\beta = 1$. For this case,
+
+\begin{align}
+X(t + T) = 
+\left(
+\begin{array}{cccccc}
+0 & 0 & 1 & 0 & 0 & 0 \\
+0 & 0 & 1 & 0 & 0 & 0 
+\end{array}  
+\right)
+\end{align}
+
+and the effect is therefore 
+\begin{align}
+y = 
+\left(
+\begin{array}{cccccc}
+0 & 0 & 0 & 2 & 0 & 0
+\end{array}  
+\right)
+\end{align}
+
+We must consider the effects at the same time to accurately calculate the likelihood.
+
+As shown above, if the impulses are not independent, then we need to consider time shifts $\tau$ at the same time.
+Therefore, we treat the problem of finding the best time shifts as a discrete optimisation problem.
+
+For the optimisation step we utilise two assumptions to complete the optimisation.
+Firstly, smaller shifts are more likely than larger shifts (proportionate to the standard deviation of the T distribution).
+This means that we should explore the space of smaller shifts more often than larger shifts.
+Secondly, impulses close to each other are more likely to be dependent on each other,  than impulses further away.
+
+Therefore, we initialize the optimisation algorithm with all parameters set to the mean of the T distribution (e.g.
+zero for the zero mean gaussian).
+Then we randomly select a small number of impulses, the exact number could be specified manually as a hyperparameter, or inferred from the data by looking at the density of impulses in time.
+These impulses are randomly changed to a value drawn for the T distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
+If the proposal likelihood is higher than the current maximum likelihood, then we update our initial estimate to the current best guess.
+This procedure is done in a loop, with the best guess of the set of T improving over time.
+
+Lastly, we optimise over the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$, iteratively alternating between the time shift optimisation and the parameter optimisation.
+For the parameter optimisation, typical methods can be used such as gradient descent, genetic algorithms or annealing.
+In our implementation we have used the differential evolution algorithm from the scipiy.optimize library (REF) in python, due to its ability to handle noisy objective functions.
+In order to improve convergence, we also standardize all input variables to the range 0-1.
+
+Through optimisation, we find the best fit for the values of $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$.
+The accuracy of the final parameter estimate is relative to the ratio of the $y$-axis error and the effect size $\beta X(t)$.
+As $\sigma_{\varepsilon}/(\beta X(t))$ tends to infinity, the estimated value of $\sigma_{\tau}$ tends to zero.
+In the event that $\sigma_{\tau}=0$, the model becomes a ‘fixed lag’ model, and for the discrete gaussian case the parameter estimate $\beta$ tends to the standard linear regression coefficient.
+On the other hand as $\sigma_{\varepsilon}/(\beta X(t))$ tends to zero, the estimated value of B(?) tends to the true value.
+Therefore, the method provides no guarantee on recovering the exact time shifts, instead only that we obtain an estimate equal to or better than standard regression.
+
+\section{Limitations}
+
+\section{Experiments}
+
+\section{Future Work}
+
+\section{Conclusion}
+
+We have proposed a form of regression analysis for dealing with stochastic time delay problems.
+We have shown the feasibility of the approach and its performance on simulated data.
+The approach allows for better estimation and prediction when the input is affected by noise in the time domain.
+To our knowledge, the method is novel for this class of problem.
 

### 2021-02-13 experiments/waves.py
Experimented with random walk process for time shifts on 'wavelike' inputs.

@@ -1,0 +1,144 @@
+import pandas as pd
+import numpy as np
+import os
+import matplotlib.pyplot as plt
+from scipy.signal import resample
+import time
+
+
+os.chdir("/Users/aaronpickering/Desktop/projects/cronus/")
+
+# Custom imports
+from src.model.linearObjective import linearTauSolver
+from src.model.linearTVSR import linearTVSRModel
+import src.support as src
+
+
+pi = 3.1415
+
+
+t = np.linspace(0,20*pi, num=100)
+
+#lynx = pd.read_csv("./experiments/lynx.csv")
+#lynx.columns = ["year","value"]
+
+#plt.plot(lynx['value'])
+
+#y = np.array(lynx['value'], dtype="float64")
+# Standardise
+#y = y/np.max(y)
+#t = np.linspace(0,19*pi, num=114)
+
+#x = np.sin(t)
+#x[0] = 0.0001
+
+
+x = np.sin(t)
+x[0]=0.0001
+phase = 0.05*t
+y = 2*np.sin(t + phase)
+y[0] = 0.0001
+
+plt.plot(x)
+plt.plot(y)
+
+
+# Generate a vector with resample freq x 10
+r = np.linspace(0, 20*np.pi, 1000)
+
+# Interpolate the signal and upsample * 10
+x_r = np.interp(r, t, x)
+y_r = np.interp(r, t, y)
+
+na_pos = np.arange(0,100)
+non_na_pos = np.arange(0,10)*10
+# add the last value - makes interpolation better
+non_na_pos = np.append(non_na_pos, 99)
+na_pos = na_pos[~np.isin(na_pos, non_na_pos)]
+
+
+# Convert to matrix form
+XR = src.decompose_vector(x_r)
+XR[na_pos,:] = np.nan
+
+
+# Generate the shift sequence - assume the that the shifts are defined by some random walk sequence
+# With gaussian step sequences
+min_res = -10000000
+min_shifts = np.isnan
+min_sd = np.nan
+min_a = np.nan
+for i in range(0,100000):
+    sd = np.random.uniform()
+    a = np.random.randn()*1.5
+    tsd = np.random.uniform()*2
+    print(i)
+    jumps = np.round(np.random.randn(XR.shape[0])*tsd)
+    jumps = np.array(jumps, dtype="int")
+    C = 0
+    
+    shift = [C]
+    for j in range(1,XR.shape[0]):
[…]

### 2021-02-13 documentation/article/article.tex
Test change x -> X
Added latex text for 'Limitations' section.

@@
 % TODO: Rewrite concrete expression of support...is it on x ot t?
+-- Aaron: f(x(t)).
+Consider x(t) = 0.1 for t is odd, 1 for t is even.
+If f(x(t)) = 0 for t is odd, 1 for t is even then the series is still valid for our method.
 \textcolor{red}{For clarity, as per convention, we use the nomenclature ‘support’ for the elements $f(x(t)) \neq 0$}.
-In addition, the series $x(t)$ should be sparse, in that it contains a low density of ‘non-zero’ coordinates.
+The size of the support of f(x(t)) should be small relative to the total vector length.
+We utilise the terminology 'impulse' for an individual element of the support set.
 
@@
 \begin{align}
-x(t) = 
+X(t) = 
 \left(
@@
 Firstly, we define a parametric form and some initial parameters to be estimated for the function $f$.
-For example, let's take the simple univariate linear model where the $f(x(t))$ is parameterised by $\beta$, $\sigma_{\varepsilon}$(error standard deviation)  and $\sigma_{\tau}$ (time error standard deviation).
+For simplicity, let's take the univariate linear model where the $f(x(t))$ is parameterised by $\beta$, $\sigma_{\varepsilon}$(error standard deviation)  and $\sigma_{\tau}$ (time error standard deviation).
 First, we initialise some starting values for each of these parameters.
 
-Now we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$.
+Now, we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$.
 It stands to reason that the best possible time shift would be one that is not too far away from the observed impulse and also gives the best possible prediction.
-In this example, we can get the prediction y by simply multiplying the shifted value by its parameter $\beta$.
-From there we can calculate the likelihood estimate for time shift plus the prediction error.
+In this example, we can calculate the prediction 'y' by simply multiplying the shifted value by its parameter $\beta$.
+From there, the likelihood estimate is derived from the time shift and the prediction error.
 
-In principle, we can then try a number of values of $\tau$ (i.e. optimise) until we maximise the likelihood for this impulse.
+In principle, we can iterate over a number of values of $\tau$ (i.e. optimise) until we maximise the likelihood for the specific impulse.
 
-However, we must also consider that the impulses in $X(t)$ are not independent from each other. After shifting, it's possible that two or more effects can occur simultaneously.
+However, we must also consider that the impulses in $X(t)$ are not independent from each other. After shifting, it's possible that two or more effects occur simultaneously.
 
-This could be particularly problematic if there are multiple impulses within a short period of time, or the impulses have a distributed effect over multiple time points.
+This is particularly problematic if there are multiple impulses within a short period of time, or impulses have a distributed effect over multiple time points.
 
@@
 
-We must consider the effects at the same time to accurately calculate the likelihood.
+To accurately calculate the likelihood, we must optimise the time shifts simultaneously.
 
-As shown above, if the impulses are not independent, then we need to consider time shifts $\tau$ at the same time.
-Therefore, we treat the problem of finding the best time shifts as a discrete optimisation problem.
+As shown above, if the impulses are not independent we need to optimise the time shifts $\tau$ simultaneously.
+Therefore, we treat the problem of finding the best set of time shifts as a discrete optimisation problem.
 
-For the optimisation step we utilise two assumptions to complete the optimisation.
+For the optimisation step we utilise two assumptions.
 Firstly, smaller shifts are more likely than larger shifts (proportionate to the standard deviation of the T distribution).
-This means that we should explore the space of smaller shifts more often than larger shifts.
-Secondly, impulses close to each other are more likely to be dependent on each other,  than impulses further away.
+The algorithm should explore the space of smaller shifts more often than larger shifts.
+Secondly, impulses close to each other are more likely to be dependent than impulses further away.
 
@@
 zero for the zero mean gaussian).
-Then we randomly select a small number of impulses, the exact number could be specified manually as a hyperparameter, or inferred from the data by looking at the density of impulses in time.
-These impulses are randomly changed to a value drawn for the T distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
-If the proposal likelihood is higher than the current maximum likelihood, then we update our initial estimate to the current best guess.
-This procedure is done in a loop, with the best guess of the set of T improving over time.
+Next, we randomly select a small number of impulses, with the value treated as a hyperparameter.
+The impulses are then randomly changed to a value drawn for the T distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
+If the proposal likelihood is higher than the current maximum likelihood, we update our initial estimate to the current best guess.
+This procedure is done in a loop with the best guess of the set of T improving over time.
 
 Lastly, we optimise over the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$, iteratively alternating between the time shift optimisation and the parameter optimisation.
-For the parameter optimisation, typical methods can be used such as gradient descent, genetic algorithms or annealing.
-In our implementation we have used the differential evolution algorithm from the scipiy.optimize library (REF) in python, due to its ability to handle noisy objective functions.
+For the parameter optimisation, typical methods such as gradient descent, genetic algorithms or annealing can be used.
+In our implementation we have used the differential evolution algorithm from scipiy.optimize (REF), because of its ability to handle noisy objective functions.
 In order to improve convergence, we also standardize all input variables to the range 0-1.
@@
 In the event that $\sigma_{\tau}=0$, the model becomes a ‘fixed lag’ model, and for the discrete gaussian case the parameter estimate $\beta$ tends to the standard linear regression coefficient.
-On the other hand as $\sigma_{\varepsilon}/(\beta X(t))$ tends to zero, the estimated value of B(?) tends to the true value.
-Therefore, the method provides no guarantee on recovering the exact time shifts, instead only that we obtain an estimate equal to or better than standard regression.
+On the other hand as $\sigma_{\varepsilon}/(\beta X(t))$ tends to zero, the coefficient estimate tends to the true value.
+Therefore, the method provides no guarantee on recovering the exact time shifts, only that the estimates are equal to or better than their OLS counterparts.
 
 \section{Limitations}
+
+As the length of $x(t)$ increases, more impulses are introduced and the size of the decomposed matrix $X(t)$ also increases.
+At some point, handling $X(t)$ becomes impractical.
+To account for this, we assume that distant impulses do not affect each other.
+Concretely, for the $ith$ and $jth$ impulse (row) in $X(t)$, separated by time $t$, the effect vectors $f(Xi*(t + tau_i))$ and $f(Xj*(t + tau_j))$ are linearly independent.
+
+Therefore, the matrix $X(t)$ is decomposed into a number of segments, where each segment is 
+composed of a smaller number of dependent impulses.
+To select the right split locations, we search the time series for locations with the lowest density of impulses.
+Assuming that we wish to split $X(t)$ into independent matrices $Un*(t)$ and $Vm*(t)$, if we choose the split locations such that $\forall_i \epsilon n and  \forall_j \epsilon_ m$, $f(Ui*(t + tau_i))$ is linearly independent from $f(Vj*(t + tau_j))$, then segments will also be independent and the procedure unique.
+In practice, it may not be possible to guarantee complete independence of every segment, in which case the results will vary.
+After segmentation, each matrix can be optimised independently, in parallel, making the inner optimisation procedure tractable for longer length sequences.
+
+Finally, a note on problem constraints.
+The likelihood estimates are dependent on the inner ($T$) optimisation procedure.
+There is no guarantee that the global maximum will be found during this process, particularly for sequences with a high density of impulses.
+
+For example, consider the example sequence
+
+\begin{align}
+    x(t) = 
+    \left(
+\begin{array}{cccccccccc}
+    0.5 & 0.6 & 0.4 & 1 & 2 & 1 &
+    0.3 & 0.2 & 0.5 & 0.8 
+\end{array}  
+\right)
+\end{align}
+
+In cases such as these, the estimated likelihood is likely to be close to but not exactly the same as the real value.
+As the density (in time) of impulses increases, the accuracy of the estimates decrease.
+Therefore, this model is most appropriate for sparse time series inputs.
 
@@
 
-We have proposed a form of regression analysis for dealing with stochastic time delay problems.
+We have proposed a form of regression analysis for handling stochastic time delay problems.
 We have shown the feasibility of the approach and its performance on simulated data.

### 2021-02-13 documentation/article/article.tex
Improved some notation and cleaned some typos.

@@
 \begin{align}
-y(t)= f(X(t + T)) + \varepsilon
+y(t)= f(X(t + \mathcal{T})) + \varepsilon
 \end{align}
@@
 To begin with, the values of each individual time shift ($\tau$) are not known.
-In addition, the size of prediction error $\varepsilon$ can only be determined if each tau is known (because for each time shift there is a different prediction and prediction error).
-Therefore, our algorithm finds the best estimate of the individual time shifts in an inner optimisation, then iteratively searches for the best estimate of $f$ in an outer optimisation step.
+In addition, prediction error distribution $\varepsilon$ can only be determined if each $\tau$ is known (because for each time shift there is a different prediction and prediction error).
+Therefore, our algorithm finds the optimum set of time shifts  in an inner optimisation, while iteratively searching for the optimum parameters of $f$ in an outer optimisation step.
 
@@
 
-with $T= (1, 0)$ and $\beta = 1$. For this case,
+with $\mathcal{T}= (1, 0)$ and $\beta = 1$. For this case,
 
 \begin{align}
-X(t + T) = 
+X(t + \mathcal{T}) = 
 \left(
@@
 
-As shown above, if the impulses are not independent we need to optimise the time shifts $\tau$ simultaneously.
 Therefore, we treat the problem of finding the best set of time shifts as a discrete optimisation problem.
@@
 For the optimisation step we utilise two assumptions.
-Firstly, smaller shifts are more likely than larger shifts (proportionate to the standard deviation of the T distribution).
+Firstly, smaller shifts are more likely than larger shifts (proportionate to the standard deviation of the $\mathcal{T}$ distribution).
 The algorithm should explore the space of smaller shifts more often than larger shifts.
@@
 
-Therefore, we initialize the optimisation algorithm with all parameters set to the mean of the T distribution (e.g.
+We initialize the optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution (e.g.
 zero for the zero mean gaussian).
 Next, we randomly select a small number of impulses, with the value treated as a hyperparameter.
-The impulses are then randomly changed to a value drawn for the T distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
+The shifts are then randomly changed to a value drawn for the T distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
 If the proposal likelihood is higher than the current maximum likelihood, we update our initial estimate to the current best guess.
@@
 \section{Limitations}
-
+%TO-DO : Improve the formulas
 As the length of $x(t)$ increases, more impulses are introduced and the size of the decomposed matrix $X(t)$ also increases.
@@
 Assuming that we wish to split $X(t)$ into independent matrices $Un*(t)$ and $Vm*(t)$, if we choose the split locations such that $\forall_i \epsilon n and  \forall_j \epsilon_ m$, $f(Ui*(t + tau_i))$ is linearly independent from $f(Vj*(t + tau_j))$, then segments will also be independent and the procedure unique.
-In practice, it may not be possible to guarantee complete independence of every segment, in which case the results will vary.
+In practice, it may not be possible to guarantee complete independence of every segment, in which case results will vary.
 After segmentation, each matrix can be optimised independently, in parallel, making the inner optimisation procedure tractable for longer length sequences.
@@
 Finally, a note on problem constraints.
-The likelihood estimates are dependent on the inner ($T$) optimisation procedure.
+The likelihood estimates are dependent on the inner ($\mathcal{T}$) optimisation procedure.
 There is no guarantee that the global maximum will be found during this process, particularly for sequences with a high density of impulses.

### 2021-02-13 documentation/article/article.tex
Began on the experiments section.

@@
 \usepackage{xcolor}
+\usepackage{hyperref}
 
@@
 
-In cases such as these, the estimated likelihood is likely to be close to but not exactly the same as the real value.
+In such cases, the estimated likelihood is likely to be close to but not exactly the same as the real value.
 As the density (in time) of impulses increases, the accuracy of the estimates decrease.
@@
 \section{Experiments}
+
+The following section demonstrates a simulated univariate example. Full code to reproduce the example can be found via \href{https://github.com/aaron1rcl/tvs_regression/}{TVS Regression on Github}.
+
+Figure 2. shows the simulated time series.
+
+The input signal has 17 non-zero impulses which have been drawn from the standard normal distribution. The system is 'off' when $x(t) =0$. In other words, when $x(t) = 0$, $f(x(t)) = 0$.
+
+The true shift distribution $\mathcal{T}$ is given by $\mathcal{T}$ $~ N(0, 2)$. There is one value of $\tau$ for each of the 17 impulses.
+
+The green line represents the shifted series, corresponding to the time at which the effect occurs. The blue line is the output ‘y’.
+
+The values for the parameters were arbitrarily selected. The output y is defined by the following equation:
+
+\begin{align}
+    y = 2X(t + \mathcal{T}) + 6.5
+\end{align}
+
+\begin{center}
+\begin{figure}
+\includegraphics[totalheight=6cm]{images/simulated_example.png}
+\caption{Simulated Example Data}
+\label{fig:verticalcell}
+\end{figure}
+\end{center}
+
+A histogram of the actual distribution of $\mathcal{T}$ is shown in figure 3.
+
+\begin{center}
+\begin{figure}
+\includegraphics[totalheight=6cm]{images/real_shifts.png}
+\caption{Real Shift Distribution $\mathcal{T}$}
+\label{fig:verticalcell}
+\end{figure}
+\end{center}
+
+After fitting the model, we obtain the following results.
+Figures 4 and 5 show the model fit (denoted TVS) and a comparison with standard OLS. Figures 6 and 7 fit convergence and scatterplot respectively.
+
+\begin{center}
+\begin{figure}
+\includegraphics[totalheight=6cm]{images/tvs_fit.png}
+\caption{TVS Model Fit $\mathcal{T}$}
+\label{fig:verticalcell}
+\end{figure}
+\end{center}
+
+The true set of $\mathcal{T}$ is:
+
+Compared to the estimated $\mathcal{T}$:
 

### 2021-02-14 documentation/article/article.tex
Added experiment, future work, conclusion

@@
 \usepackage{hyperref}
+\usepackage{tabularx}
 
@@
 To account for this, we assume that distant impulses do not affect each other.
-Concretely, for the $ith$ and $jth$ impulse (row) in $X(t)$, separated by time $t$, the effect vectors $f(Xi*(t + tau_i))$ and $f(Xj*(t + tau_j))$ are linearly independent.
+Concretely, for the $ith$ and $j_th$ impulse (row) in $X(t)$, separated by time $t$, the effect vectors $f(X_i(t + \tau_i))$ and $f(X_j(t + \tau_j))$ are linearly independent.
 
@@
 composed of a smaller number of dependent impulses.
-To select the right split locations, we search the time series for locations with the lowest density of impulses.
-Assuming that we wish to split $X(t)$ into independent matrices $Un*(t)$ and $Vm*(t)$, if we choose the split locations such that $\forall_i \epsilon n and  \forall_j \epsilon_ m$, $f(Ui*(t + tau_i))$ is linearly independent from $f(Vj*(t + tau_j))$, then segments will also be independent and the procedure unique.
-In practice, it may not be possible to guarantee complete independence of every segment, in which case results will vary.
-After segmentation, each matrix can be optimised independently, in parallel, making the inner optimisation procedure tractable for longer length sequences.
+To select the right split locations, we search the time series for locations with the lowest impulse density.
+Given that we wish to split $X(t)$ into independent matrices $U_n(t)$ and $V_m(t)$,
+if we choose the split locations such that $\forall_i$ $\epsilon$ $n$ and  $\forall_j$ $\epsilon$ $m$, 
+$f(U_i(t + \tau_i))$ is linearly independent from $f(V_j(t + \tau_j))$, 
+then segments will also be independent and the procedure unique. 
+In practice, it may not be possible to guarantee complete independence of every segment, 
+in which case results may vary. After segmentation, each matrix can be optimised independently, 
+in parallel, making the inner optimisation procedure tractable for longer length sequences.
 
@@
 
+where each impulse can shift between -2 and +2 positions. Ignoring shifts beyond the edge of the vector, the solutions space of $\mathcal{T}$ is $5^t$ combinations.  
 In such cases, the estimated likelihood is likely to be close to but not exactly the same as the real value.
@@
 
-\section{Experiments}
+\section{Experiment}
 
@@
 
+\begin{align}
+    \mathcal{T} = 
+    \left(
+    \begin{array}{ccc}
+    -1 & 1 & 0 
+    \end{array}  
+    \right)
+    \end{align}
+
 Compared to the estimated $\mathcal{T}$:
+
+\begin{align}
+    \mathcal{T}_{est} = 
+    \left(
+    \begin{array}{ccc}
+    -1 & 1 & 0 
+    \end{array}  
+    \right)
+    \end{align}
+
+%To-do - write in the actual T distributions above. Not sure how to fix the hbox problem.
+
+%To-do properly format the table
+\begin{tabular}{c|c|c|c|}
+\hline
+Parameter & TVS Regression & OLS Regression & True Value \\
+\hline
+$\beta$ & 1.96 & 0.26 & 2  \\
+\hline
+$Intercept$ & 6.48 & 6.62 & 6.5 \\
+\hline
+$\sigma_t$ & 1.51 & $N/A$ & 1.47 \\
+\hline
+$\sigma_e$ & 0.26 & 1.03 & 0.2 \\
+\end{tabular}
+
+As shown above, the estimated values for $\beta$ and $\sigma_e$ are significantly improved by taking into account the stochastic time delay noise.
+The estimated $\beta$ using OLS regression is 0.25, compared to the true value 2.
+Even a small amount of noise in the time axis causes attenuation,
+limiting the usefulness of OLS for these problems.
+In comparison, the algorithm described herein estimates a $\beta$ of 1.96, much closer to the true value.
+
+While the example is based on simulated data, we believe that the experiment demonstrates clear potential for improvement in the modelling of real world systems with stochastic time delay noise.
 
@@
 
+In the univariate case, the utility of the described method is limited.
+However, we propose three extensions as future work.
+The first is to extend the method to multivariate regression.
+We believe that the extension can be built on the same fundamental ideas presented in this document.
+Next, the model could be extended to include distributed lag structures.
+A distributed lag structure is where past values of the impulse influence future values of the output.
+Finally, in relating $x(t)$ to $f(x(t))$, the function f could be extended to the modelling of non-linear forms.
+
+Through these extensions we can begin to tackle a larger number of practical problems.
+
+In addition, by incorporating these elements we could begin to explore alternative forms of time series prediction.
+Essentially, existing prediction mechanisms predict the average across time shifts.
+In some situations, the predictions resulting from these models could be considered nonsensical.
+
+For example, let us consider a regression model to predict how much coffee someone drinks between 9am and 9:30am, at work.
+They set off every day in their car at exactly 8:30am and it takes about 30 minutes (+- $\tau$) to get to work.
+The time shift $\tau$ could be caused by something as simple as variations in traffic.
+When they get to work, they always have a coffee.
+A typical regression might predict $0.5$ coffees between 8:30 am and 9am and $0.5$ coffees between 9am and 9:30am.
+Alternative prediction forms such as: 'When they get to work, they will have 1 coffee' might be more useful.
+
 \section{Conclusion}
 
-We have proposed a form of regression analysis for handling stochastic time delay problems.
-We have shown the feasibility of the approach and its performance on simulated data.
-The approach allows for better estimation and prediction when the input is affected by noise in the time domain.
+We have proposed a form of regression analysis suited to the modelling of stochastic time delay problems.
+In addition, we have shown the feasibility of the approach and its performance on simulated data.
+Our approach allows for consistently improved estimation and prediction when the input is affected by noise in the time domain.
 To our knowledge, the method is novel for this class of problem.

### 2021-02-14 documentation/article/article.tex
Rewording.
Fix typo
Wording change

@@
 
-where $X(t + T)$ is the time shifted series and $X(t)$ is the observed input.\\
+where $X(t + \mathcal{T})$ is the matrix of time shifted impulses and $X(t)$ is the observed input.\\
 
@@
 However, we propose three extensions as future work.
-The first is to extend the method to multivariate regression.
+The first is to extend the method to multiple regression.
 We believe that the extension can be built on the same fundamental ideas presented in this document.
@@
 In addition, by incorporating these elements we could begin to explore alternative forms of time series prediction.
-Essentially, existing prediction mechanisms predict the average across time shifts.
-In some situations, the predictions resulting from these models could be considered nonsensical.
+Essentially, existing prediction mechanisms predict the average across time shifts.However in some situations, the predictions could be considered nonsensical.
 

### 2021-02-14 documentation/article/article.tex
Typo fix

@@
 In addition, by incorporating these elements we could begin to explore alternative forms of time series prediction.
-Essentially, existing prediction mechanisms predict the average across time shifts.However in some situations, the predictions could be considered nonsensical.
+Essentially, existing prediction mechanisms predict the average across time shifts.However, in some situations the predictions could be considered nonsensical.
 

### 2021-02-14 documentation/article/article.tex
Title change
add images

@@
 
-\title{Modelling Stochastic Time Delay Structures for Regression}
+\title{Modelling Stochastic Time Delay for Regression}
 \author{Aaron Pickering}
@@
 
-The following section demonstrates a simulated univariate example. Full code to reproduce the example can be found via \href{https://github.com/aaron1rcl/tvs_regression/}{TVS Regression on Github}.
+The following section demonstrates a simulated univariate example. Full code to reproduce the example can be found via \href{https://github.com/aaron1rcl/tvs_regression/}{TVS Regression on GitHub}.
 
-Figure 2. shows the simulated time series.
+Figure \ref{fig:simulated_example} shows the simulated time series.
 
@@
 \begin{figure}
-\includegraphics[totalheight=6cm]{images/simulated_example.png}
+\includegraphics[scale=0.5]{images/simulated_example.png}
 \caption{Simulated Example Data}
-\label{fig:verticalcell}
+\label{fig:simulated_example}
 \end{figure}
@@
 
-A histogram of the actual distribution of $\mathcal{T}$ is shown in figure 3.
+A histogram of the actual distribution of $\mathcal{T}$ is shown in Figure \ref{fig:real_shifts}.
 
@@
 \begin{figure}
-\includegraphics[totalheight=6cm]{images/real_shifts.png}
-\caption{Real Shift Distribution $\mathcal{T}$}
-\label{fig:verticalcell}
+\includegraphics[scale=0.5]{images/real_shifts.png}
+\caption{Real Shift Distribution $\mathcal{T}$.}
+\label{fig:real_shifts}
 \end{figure}
@@
 After fitting the model, we obtain the following results.
-Figures 4 and 5 show the model fit (denoted TVS) and a comparison with standard OLS. Figures 6 and 7 fit convergence and scatterplot respectively.
+Figure \ref{fig:tvs_ols_fit} show the model fit (denoted TVS) and a comparison with standard OLS.
+Figure \ref{fig:tvs_errors} and Figure \ref{fig:convergence} show the errors distribution of the TVS fit and  convergence of the TVS model respectively.
 
@@
 \begin{figure}
-\includegraphics[totalheight=6cm]{images/tvs_fit.png}
-\caption{TVS Model Fit $\mathcal{T}$}
-\label{fig:verticalcell}
+\includegraphics[scale=0.6]{images/tvs_fit.png} 
+\includegraphics[scale=0.3]{images/ols_fit.png} 
+\includegraphics[scale=0.3]{images/scatterplot.png} 
+\caption{TVS vs OLS model fit.}
+\label{fig:tvs_ols_fit}
+\end{figure}
+\end{center}
+
+\begin{center}
+\begin{figure}
+\includegraphics[scale=0.5]{images/tvs_errors.png}
+\caption{Errors...}
+\label{fig:tvs_errors}
+\end{figure}
+\end{center}
+
+\begin{center}
+\begin{figure}
+\includegraphics[scale=0.6]{images/convergence.png} 
+\caption{Convergence ...}
+\label{fig:convergence}
 \end{figure}

### 2021-02-15 documentation/article/article.tex
Improve the loss function formulas and wording.

@@
 \begin{align}
-y(t)= f(X(t + \mathcal{T})) + \varepsilon
+y(t)= f(1^TX(t + \mathcal{T})) + \varepsilon
 \end{align}
 
-where $X(t + \mathcal{T})$ is the matrix of time shifted impulses and $X(t)$ is the observed input.\\
+where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input and 1 is the vector (1,1,1...1).\\
 
-In order to find the best estimate of $f$, we would like to find the function $f$ which maximises the joint likelihood of both Tau(?) and $\varepsilon$, the time-domain likelihood and the error likelihood respectively.
+In order to find the best estimate of $f$, we would like to find the function $f$ which maximises the joint likelihood of the time-domain shifts and the prediction error likelihood.
 
-Specifically, we maximise the log-likelihood of the following relation:
+Specifically, we maximise:
 % TODO: Clean this notation.
 \begin{align}
-\max  \left( \log \left(\mathcal{L}(\theta_{\varepsilon}, \varepsilon) \times \mathcal{L}(\theta_{\tau}, \tau)\right) \right)
+    \sum_{i=1}^{n} \log(p(y_i | X = x_i; \theta_{\varepsilon}, \tau_i)) + \sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}))
 \end{align}
@@
 where $\theta_{\varepsilon}$ and $\theta_{\tau}$ represent the parameters of the prediction error and time error respectively.
-For simplicity, assume that the tau and error distributions are independent.
-To begin with, the values of each individual time shift ($\tau$) are not known.
-In addition, prediction error distribution $\varepsilon$ can only be determined if each $\tau$ is known (because for each time shift there is a different prediction and prediction error).
-Therefore, our algorithm finds the optimum set of time shifts  in an inner optimisation, while iteratively searching for the optimum parameters of $f$ in an outer optimisation step.
+For simplicity, we assume that the time shift ($\mathcal{T}$) and error ($\varepsilon$) distributions are independent.
 
@@
 
+Before optimisation, the values of each individual time shift ($\tau$) are not known.
+In addition, the prediction errors $\varepsilon$ can only be calculated if each value of $\tau$ is available (because for each time shift there is a different prediction and hence prediction error).
+Therefore, our algorithm finds the optimum set of time shifts in an inner optimisation, while iteratively searching for the optimum parameters of $f$ in an outer optimisation.
 Firstly, we define a parametric form and some initial parameters to be estimated for the function $f$.
@@
 \begin{align}
-    y = 2X(t + \mathcal{T}) + 6.5
+    y = 2(1^TX(t + \mathcal{T})) + 6.5
 \end{align}

### 2021-02-15 documentation/article/article.tex
Improved some formulas, including the loss function and in the algorithm section.

@@
 discrete gaussian, poisson etc).
-For the discrete gaussian kernel $\mathcal{T}\sim T(0, \sigma)$ or for the poisson distribution $\mathcal{T}\sim Pois(\lambda)$ and $\varepsilon \sim N(0, \sigma_{\varepsilon})$.
+For the discrete gaussian kernel $\mathcal{T}\sim T(\mu, \sigma)$ or for the poisson distribution $\mathcal{T}\sim Pois(\lambda)$ and $\varepsilon \sim N(\mu, \sigma_{\varepsilon})$.
 
@@
 
-where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input and 1 is the vector (1,1,1...1).\\
+where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input and 1 is a vector (1,1,1...1).\\
 
-In order to find the best estimate of $f$, we would like to find the function $f$ which maximises the joint likelihood of the time-domain shifts and the prediction error likelihood.
+In order to find the best estimate of $f$, we would like to find the function $f$ which maximises the joint likelihood of the time-domain shifts and the prediction residuals.
 
@@
 \begin{align}
-    \sum_{i=1}^{n} \log(p(y_i | X = x_i; \theta_{\varepsilon}, \tau_i)) + \sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}))
+    \sum_{i=1}^{n} \log(p(y_i | X_{\tau i} = x_i; \theta_{\mathcal{f}}, \tau_i)) + \sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}))
 \end{align}
 
-where $\theta_{\varepsilon}$ and $\theta_{\tau}$ represent the parameters of the prediction error and time error respectively.
-For simplicity, we assume that the time shift ($\mathcal{T}$) and error ($\varepsilon$) distributions are independent.
+where $\theta_{f}$ and $\theta_{\mathcal{T}}$ represent the parameters of the model $f$ and time shifts $\mathcal{T}$.
+For simplicity, we assume that the time shift distribution and error distributions are independent.
 
@@
 
-In principle, we can iterate over a number of values of $\tau$ (i.e. optimise) until we maximise the likelihood for the specific impulse.
+Finally, we iterate over a number of values of $\tau$ optimising until we maximise the likelihood for the specific impulse.
 
@@
 Through optimisation, we find the best fit for the values of $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$.
-The accuracy of the final parameter estimate is relative to the ratio of the $y$-axis error and the effect size $\beta X(t)$.
-As $\sigma_{\varepsilon}/(\beta X(t))$ tends to infinity, the estimated value of $\sigma_{\tau}$ tends to zero.
+The accuracy of the final parameter estimate is relative to the ratio of the $y$-axis error and the effect size $f(x(t))$.
+As $\sigma_{\varepsilon}/f(x(t))$ $\rightarrow$ $\infty$, $\sigma_{\tau}$ $\rightarrow$ $0$.
 In the event that $\sigma_{\tau}=0$, the model becomes a ‘fixed lag’ model, and for the discrete gaussian case the parameter estimate $\beta$ tends to the standard linear regression coefficient.
-On the other hand as $\sigma_{\varepsilon}/(\beta X(t))$ tends to zero, the coefficient estimate tends to the true value.
-Therefore, the method provides no guarantee on recovering the exact time shifts, only that the estimates are equal to or better than their OLS counterparts.
+On the other hand as $\sigma_{\varepsilon}/f(x(t))$ tends to zero, $\beta x(t)$ $\rightarrow f(x(t))$.
+The method provides no guarantee on recovering the exact time shifts, only that the estimates are equal to or better than their OLS counterparts.
 
 \section{Limitations}
-%TO-DO : Improve the formulas
 As the length of $x(t)$ increases, more impulses are introduced and the size of the decomposed matrix $X(t)$ also increases.

### 2021-02-15 documentation/article/article.tex
Improve wording and formulas.

@@
 
-Consider the simple example where the effect of the input is $1$, as shown in the image below.
+Consider the simple example in Figure 1 where the effect of the input is $1$.
 The observed input is given by the red line, the blue line is when the effect actually occurs.
@@
 Obviously, this isn't ideal, we want the parameter estimates to be as close to the real values as possible, regardless of any noise in the lag structures.
-One can mitigate the problem via time aggregations, however in complicated multivariate cases, this is just not feasible.
+One can mitigate the problem via time aggregations, however in complicated cases with multiple factors, this is just not feasible.
 
@@
 
-Our algorithm then attempts to find the maximum likelihood estimate for the following equation:
+We then model:
 
@@
 
-where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input and 1 is a vector (1,1,1...1).\\
+where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input, $1$ is the vector $(1,1,1...1)$ and $\epsilon$ is a noise term.\\
 
@@
 
-where $\theta_{f}$ and $\theta_{\mathcal{T}}$ represent the parameters of the model $f$ and time shifts $\mathcal{T}$.
+where $\theta_{f}$ and $\theta_{\mathcal{T}}$ represent the parameters of the model $f$ and time shifts $\mathcal{T}$ respectively.
 For simplicity, we assume that the time shift distribution and error distributions are independent.
@@
 In addition, the prediction errors $\varepsilon$ can only be calculated if each value of $\tau$ is available (because for each time shift there is a different prediction and hence prediction error).
-Therefore, our algorithm finds the optimum set of time shifts in an inner optimisation, while iteratively searching for the optimum parameters of $f$ in an outer optimisation.
+Therefore, our algorithm finds the optimum set of time shifts in an inner optimisation, while iteratively searching for the optimum parameters of $f$ in an outer optimisation loop.
 Firstly, we define a parametric form and some initial parameters to be estimated for the function $f$.
@@
 The shifts are then randomly changed to a value drawn for the T distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
-If the proposal likelihood is higher than the current maximum likelihood, we update our initial estimate to the current best guess.
+If the proposal likelihood is higher than the current maximum likelihood, we update our estimate.
 This procedure is done in a loop with the best guess of the set of T improving over time.
@@
 In our implementation we have used the differential evolution algorithm from scipiy.optimize (REF), because of its ability to handle noisy objective functions.
-In order to improve convergence, we also standardize all input variables to the range 0-1.
+In order to improve convergence, we also standardize all input variables to the range (0,1).
 
@@
 To account for this, we assume that distant impulses do not affect each other.
-Concretely, for the $ith$ and $j_th$ impulse (row) in $X(t)$, separated by time $t$, the effect vectors $f(X_i(t + \tau_i))$ and $f(X_j(t + \tau_j))$ are linearly independent.
+Concretely, for the $i_{th}$ and $j_{th}$ impulse (row) in $X(t)$, separated by time $t \rightarrow \infty$, the effect vectors $f(X_i(t + \tau_i))$ and $f(X_j(t + \tau_j))$ are linearly independent.
 
@@
 In practice, it may not be possible to guarantee complete independence of every segment, 
-in which case results may vary. After segmentation, each matrix can be optimised independently, 
+in which case results may vary. After segmentation, each matrix can be optimised independently 
 in parallel, making the inner optimisation procedure tractable for longer length sequences.
@@
 The likelihood estimates are dependent on the inner ($\mathcal{T}$) optimisation procedure.
-There is no guarantee that the global maximum will be found during this process, particularly for sequences with a high density of impulses.
+This procedure is $NP$-$hard$ and there is no gaurantee that the global maximum will be found, particularly for sequences with a high density of impulses.
 
@@
 
-The true shift distribution $\mathcal{T}$ is given by $\mathcal{T}$ $~ N(0, 2)$. There is one value of $\tau$ for each of the 17 impulses.
+The true shift distribution $\mathcal{T}$ is given by $\mathcal{T}$ $~$ $N(0, 2)$. There is one value of $\tau$ for each of the 17 impulses.
 

### 2021-03-20 documentation/article/article.tex
Improve formulas.
add math comments and improve table

@@
 \maketitle
+
+\begin{abstract}
+... 
+\end{abstract}
 
@@
 We consider the problem as analogous to the typical error-in-variables (EIV) regression.
-Ordinary regression analyses (and machine learning models) define the loss function with respect to errors in the y axis only.
+Ordinary regression analyses (and machine learning models) define the loss function with respect to errors in the $y$ axis only.
 For EIV, errors are considered in both the $y$-axis and the $x$-axis (REF).
@@
 
-Let $x: [a, b] \longrightarrow \mathbb{R}$ be a discrete time series with finite domain and we want to determine the injective, functional relationship $f$ to some other variable y.
-The series $x(t)$ must be a stationary series with a known value for the point at which $f(x(t)) = 0$.
-That is, we know at which point the input series has no effect on the output (when it’s off).
-
 % TODO: Rewrite concrete expression of support...is it on x ot t?
@@
 If f(x(t)) = 0 for t is odd, 1 for t is even then the series is still valid for our method.
-\textcolor{red}{For clarity, as per convention, we use the nomenclature ‘support’ for the elements $f(x(t)) \neq 0$}.
-The size of the support of f(x(t)) should be small relative to the total vector length.
-We utilise the terminology 'impulse' for an individual element of the support set.
+Let $x: [a, b] \longrightarrow \mathbb{R}$ be a discrete time series with finite domain (\textcolor{red}{What do we mean with finite domain?
+maybe is better to use the notation $x: \mathcal{D}_n\subset [a, b] \longrightarrow \mathbb{R}$, where $\mathcal{D}_n$ is a finite set of $n$ points}).
+We want to determine the, functional relationship $f:\mathbb{R} \longrightarrow \mathbb{R}$ to some other (target) variable $y$.
+We further impose the extra condition for $f$ to be injective.
+The series $x(t)$ must be stationary with a known value for the point at which $f(x(t)) = 0$ (\textcolor{red}{How can we know this point if we do not know $f$?}).
+That is, we know at which point the input series has no effect on the output (when it’s off).
+The size of the support of $f(x(t))$ should be small relative to the total vector length (\textcolor{red}{We need to define what do we mean here by "small"}).
+We utilise the terminology {\em impulse} for an individual element of the support of $f$.
 
 We also assume that the set of $\tau$'s, denoted $\mathcal{T}$, is not a constant, but rather a random draw from some discrete distribution (e.g.
-discrete gaussian, poisson etc).
+discrete gaussian, poisson etc) (\textcolor{red}{We have not defined in the section what "the set of $\tau$'s is"}).
 For the discrete gaussian kernel $\mathcal{T}\sim T(\mu, \sigma)$ or for the poisson distribution $\mathcal{T}\sim Pois(\lambda)$ and $\varepsilon \sim N(\mu, \sigma_{\varepsilon})$.
@@
 
-where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input, $1$ is the vector $(1,1,1...1)$ and $\epsilon$ is a noise term.\\
+where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input, $1$ is the vector $(1,1,1...1)$ and $\varepsilon$ is a noise term.\\
 
 In order to find the best estimate of $f$, we would like to find the function $f$ which maximises the joint likelihood of the time-domain shifts and the prediction residuals.
-
 Specifically, we maximise:
@@
 \begin{align}
-    \sum_{i=1}^{n} \log(p(y_i | X_{\tau i} = x_i; \theta_{\mathcal{f}}, \tau_i)) + \sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}))
+    \sum_{i=1}^{n} \log(p(y_i | X_{\tau i} = x_i; \theta_{f}, \tau_i)) + \sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}))
 \end{align}
@@
 Now, we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$.
-It stands to reason that the best possible time shift would be one that is not too far away from the observed impulse and also gives the best possible prediction.
-In this example, we can calculate the prediction 'y' by simply multiplying the shifted value by its parameter $\beta$.
+It stands to reason that the best possible time shift would be one that is not too far away from the observed impulse and also gives the best possible prediction (\textcolor{red}{we need to try to formalize what "too far" means concretely}).
+In this example, we can calculate the prediction $y$ by simply multiplying the shifted value by its parameter $\beta$.
 From there, the likelihood estimate is derived from the time shift and the prediction error.
@@
 To accurately calculate the likelihood, we must optimise the time shifts simultaneously.
+Therefore, we treat the problem of finding the best set of time shifts as a discrete optimisation problem.
+For the optimisation step we utilise two assumptions.
+\begin{enumerate}
+\item Firstly, smaller shifts are more likely than larger shifts (proportionate to the standard deviation of the $\mathcal{T}$ distribution).
+The algorithm should explore the space of smaller shifts more often than larger shifts.
+\item Secondly, impulses close to each other are more likely to be dependent than impulses further away. 
+\end{enumerate}
 
-Therefore, we treat the problem of finding the best set of time shifts as a discrete optimisation problem.
-
-For the optimisation step we utilise two assumptions.
-Firstly, smaller shifts are more likely than larger shifts (proportionate to the standard deviation of the $\mathcal{T}$ distribution).
-The algorithm should explore the space of smaller shifts more often than larger shifts.
-Secondly, impulses close to each other are more likely to be dependent than impulses further away.
-
-We initialize the optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution (e.g.
+(\textcolor{red}{For the description of the algorithm we might want to use an "enumerate" latex block as above to make ste steps and dependencies clear}) We initialize the optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution (e.g.
 zero for the zero mean gaussian).
 Next, we randomly select a small number of impulses, with the value treated as a hyperparameter.
-The shifts are then randomly changed to a value drawn for the T distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
+The shifts are then randomly changed to a value drawn for the T (\textcolor{red}{is this $\mathcal{T}$?})distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
 If the proposal likelihood is higher than the current maximum likelihood, we update our estimate.
@@
 The accuracy of the final parameter estimate is relative to the ratio of the $y$-axis error and the effect size $f(x(t))$.
-As $\sigma_{\varepsilon}/f(x(t))$ $\rightarrow$ $\infty$, $\sigma_{\tau}$ $\rightarrow$ $0$.
+As $\sigma_{\varepsilon}/f(x(t))$ $\rightarrow$ $\infty$, $\sigma_{\tau}$ $\rightarrow$ $0$ (\textcolor{red}{we need to write this as $\lim$ to be more formal}).
 In the event that $\sigma_{\tau}=0$, the model becomes a ‘fixed lag’ model, and for the discrete gaussian case the parameter estimate $\beta$ tends to the standard linear regression coefficient.
-On the other hand as $\sigma_{\varepsilon}/f(x(t))$ tends to zero, $\beta x(t)$ $\rightarrow f(x(t))$.
+On the other hand as $\sigma_{\varepsilon}/f(x(t))$ tends to zero, $\beta x(t)$ $\rightarrow f(x(t))$ (\textcolor{red}{we need to write this as $\lim$ to be more formal}).
 The method provides no guarantee on recovering the exact time shifts, only that the estimates are equal to or better than their OLS counterparts.
@@
 To account for this, we assume that distant impulses do not affect each other.
-Concretely, for the $i_{th}$ and $j_{th}$ impulse (row) in $X(t)$, separated by time $t \rightarrow \infty$, the effect vectors $f(X_i(t + \tau_i))$ and $f(X_j(t + \tau_j))$ are linearly independent.
+Concretely, for the $i_{th}$ and $j_{th}$ impulse (row) in $X(t)$, separated by time $t \longrightarrow \infty$, the effect vectors $f(X_i(t + \tau_i))$ and $f(X_j(t + \tau_j))$ are linearly independent.
 
@@
 $f(U_i(t + \tau_i))$ is linearly independent from $f(V_j(t + \tau_j))$, 
-then segments will also be independent and the procedure unique. 
+then segments will also be independent and the procedure unique (\textcolor{red}{fix notation}). 
 In practice, it may not be possible to guarantee complete independence of every segment, 
@@
 
-The true shift distribution $\mathcal{T}$ is given by $\mathcal{T}$ $~$ $N(0, 2)$. There is one value of $\tau$ for each of the 17 impulses.
+The true shift distribution $\mathcal{T}$ is given by $\mathcal{T} \sim N(0, 2)$. There is one value of $\tau$ for each of the 17 impulses.
 
-The green line represents the shifted series, corresponding to the time at which the effect occurs. The blue line is the output ‘y’.
+The green line represents the shifted series, corresponding to the time at which the effect occurs. The blue line is the output $y$.
 
@@
 %To-do properly format the table
+\begin{center}
+\begin{table}
 \begin{tabular}{c|c|c|c|}
@@
 \end{tabular}
+\caption{ ... } 
+\label{results_table}
+\end{table}
+\end{center}
 
-As shown above, the estimated values for $\beta$ and $\sigma_e$ are significantly improved by taking into account the stochastic time delay noise.
-The estimated $\beta$ using OLS regression is 0.25, compared to the true value 2.
+As shown in Table \ref{results_table}, the estimated values for $\beta$ and $\sigma_e$ are significantly improved by taking into account the stochastic time delay noise.
+The estimated $\beta$ using OLS regression is $0.25$, compared to the true value $2$.
 Even a small amount of noise in the time axis causes attenuation,
 limiting the usefulness of OLS for these problems.
-In comparison, the algorithm described herein estimates a $\beta$ of 1.96, much closer to the true value.
+In comparison, the algorithm described herein estimates a $\beta$ of $1.96$, much closer to the true value.
 
@@
 A distributed lag structure is where past values of the impulse influence future values of the output.
-Finally, in relating $x(t)$ to $f(x(t))$, the function f could be extended to the modelling of non-linear forms.
+Finally, in relating $x(t)$ to $f(x(t))$, the function $f$ could be extended to the modelling of non-linear forms.
 
@@
 For example, let us consider a regression model to predict how much coffee someone drinks between 9am and 9:30am, at work.
-They set off every day in their car at exactly 8:30am and it takes about 30 minutes (+- $\tau$) to get to work.
+They set off every day in their car at exactly 8:30am and it takes about 30 minutes ($\pm \tau$) to get to work.
 The time shift $\tau$ could be caused by something as simple as variations in traffic.

### 2021-03-21 documentation/article/article.tex
Fixed some highlighted points. Improved notation and clarity. Formatted algorithm using 'enumerate'.

@@
 \usepackage{tabularx}
+\bibliography{references}
 
@@
 \title{Modelling Stochastic Time Delay for Regression}
-\author{Aaron Pickering}
+\author{Aaron Pickering, Dr. Juan Camilo Orduz}
 %\address{}
@@
 If f(x(t)) = 0 for t is odd, 1 for t is even then the series is still valid for our method.
-Let $x: [a, b] \longrightarrow \mathbb{R}$ be a discrete time series with finite domain (\textcolor{red}{What do we mean with finite domain?
-maybe is better to use the notation $x: \mathcal{D}_n\subset [a, b] \longrightarrow \mathbb{R}$, where $\mathcal{D}_n$ is a finite set of $n$ points}).
+Let $x: \mathcal{D}_n\subset [a, b] \longrightarrow \mathbb{R}$, where $\mathcal{D}_n$ is a finite set of $n$ points.
 We want to determine the, functional relationship $f:\mathbb{R} \longrightarrow \mathbb{R}$ to some other (target) variable $y$.
 We further impose the extra condition for $f$ to be injective.
-The series $x(t)$ must be stationary with a known value for the point at which $f(x(t)) = 0$ (\textcolor{red}{How can we know this point if we do not know $f$?}).
+The series $x(t)$ must be stationary with a known value for the point at which $f(x(t)) = 0$.
 That is, we know at which point the input series has no effect on the output (when it’s off).
-The size of the support of $f(x(t))$ should be small relative to the total vector length (\textcolor{red}{We need to define what do we mean here by "small"}).
+The size of the support of $f(x(t))$ should be small relative to the total vector length (we we will explore this constraint in more detail in the 'Limitations' section).
 We utilise the terminology {\em impulse} for an individual element of the support of $f$.
-
-We also assume that the set of $\tau$'s, denoted $\mathcal{T}$, is not a constant, but rather a random draw from some discrete distribution (e.g.
-discrete gaussian, poisson etc) (\textcolor{red}{We have not defined in the section what "the set of $\tau$'s is"}).
-For the discrete gaussian kernel $\mathcal{T}\sim T(\mu, \sigma)$ or for the poisson distribution $\mathcal{T}\sim Pois(\lambda)$ and $\varepsilon \sim N(\mu, \sigma_{\varepsilon})$.
 
@@
 
-We then model:
+Given that each non-zero impulse (row) of $X(t)$ is affected a random time delay (denoted $\tau$), we then model:
 
@@
 
-where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input, $1$ is the vector $(1,1,1...1)$ and $\varepsilon$ is a noise term.\\
+where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, 
+ $X(t)$ is the observed input, 
+ $1$ is the vector $(1,1,1...1)$,
+ $\varepsilon$ is a noise term and $\mathcal{T}$ is the set of single impulse time delays ($\tau$'s).\\
+
+We assume that $\mathcal{T}$ is not a constant, but rather a random draw from some discrete distribution (e.g.
+discrete gaussian, poisson etc).
+For the discrete gaussian kernel $\mathcal{T}\sim T(\mu, \sigma)$ or for the poisson distribution $\mathcal{T}\sim Pois(\lambda)$ and $\varepsilon \sim N(\mu, \sigma_{\varepsilon})$.
 
@@
 Now, we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$.
-It stands to reason that the best possible time shift would be one that is not too far away from the observed impulse and also gives the best possible prediction (\textcolor{red}{we need to try to formalize what "too far" means concretely}).
+It stands to reason that the best possible time shift would be one that is not too far away from the observed impulse (i.e.
+a high likelihood given some distribution) and also produces the best possible prediction.
 In this example, we can calculate the prediction $y$ by simply multiplying the shifted value by its parameter $\beta$.
@@
 
-(\textcolor{red}{For the description of the algorithm we might want to use an "enumerate" latex block as above to make ste steps and dependencies clear}) We initialize the optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution (e.g.
-zero for the zero mean gaussian).
-Next, we randomly select a small number of impulses, with the value treated as a hyperparameter.
-The shifts are then randomly changed to a value drawn for the T (\textcolor{red}{is this $\mathcal{T}$?})distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
-If the proposal likelihood is higher than the current maximum likelihood, we update our estimate.
-This procedure is done in a loop with the best guess of the set of T improving over time.
+Accordingly, the optimisation procedure is:
+\vspace{5mm} %5mm vertical space
 
-Lastly, we optimise over the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$, iteratively alternating between the time shift optimisation and the parameter optimisation.
-For the parameter optimisation, typical methods such as gradient descent, genetic algorithms or annealing can be used.
-In our implementation we have used the differential evolution algorithm from scipiy.optimize (REF), because of its ability to handle noisy objective functions.
+\begin{enumerate}
+    \item [(1)] Initialise the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$ ($\theta_{f}$) for the function $f$
+    \item [(2)] Find the $\mathcal{T}$ which maximises the likelihood for the given set of parameters:
+        \begin{enumerate} 
+            \item[(i)] Initialize the optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution (e.g. zero for the zero mean gaussian). 
+            \item[(ii)] Next, randomly select a small number of impulses, with the value treated as a hyperparameter.
+The time shifts are then randomly changed to a value drawn for the $\mathcal{T}$ distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
+            \item[(iii)] If the proposal likelihood is higher than the current maximum likelihood, we update our estimate.
+            \item [(iv)] Return to (i) and repeat, with the best estimate of the set of $\mathcal{T}$ improving over a number of iterations.
+The number of iterations ($n$) is also treated as a hyperparameter.
+        \end{enumerate} 
+    \item [(3)] Optimise the parameters $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$ ($\theta_{f}$).
+
+\end{enumerate}
+\vspace{5mm} %5mm vertical space
+The procedure consists of an inner optimisation to find the set of $\mathcal{T}$ and an outer optimisation to find the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$. 
+By iteratively alternating between the time shift optimisation and the parameter optimisation we are able to maximise equation (4). 
+For the outer parameter optimisation, typical methods such as gradient descent, genetic algorithms and annealing can be used.
+In our implementation, we have used the differential evolution algorithm (scipiy.optimize (REF)) because of its ability to handle noisy objective functions.
 In order to improve convergence, we also standardize all input variables to the range (0,1).
 
-Through optimisation, we find the best fit for the values of $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$.
 The accuracy of the final parameter estimate is relative to the ratio of the $y$-axis error and the effect size $f(x(t))$.
-As $\sigma_{\varepsilon}/f(x(t))$ $\rightarrow$ $\infty$, $\sigma_{\tau}$ $\rightarrow$ $0$ (\textcolor{red}{we need to write this as $\lim$ to be more formal}).
+As $\sigma_{\varepsilon}/f(x(t))$ $\rightarrow$ $\infty$ , $\sigma_{\tau}$ $\rightarrow$ $0$ (\textcolor{red}{we need to write this as $\lim$ to be more formal}).
 In the event that $\sigma_{\tau}=0$, the model becomes a ‘fixed lag’ model, and for the discrete gaussian case the parameter estimate $\beta$ tends to the standard linear regression coefficient.
@@
 
+%To-DO: Write References
+
+%\printbibliography
+
+%To-DO Further questions:
+ %- what to write in the abstract?
+ %- we havent addressed previous research that covers related problems, shall we include a section?
+ %- add references for all unbacked major statements
+ %- should we address identifiability?
+ %- examplte notebook should be fully aligned with the paper - fix
+
 \end{document}

### 2021-03-21 documentation/article/article.tex
Began the abstract.

@@
 \begin{abstract}
-... 
+Systems with stochastic time delay between input and output present a number of unique challenges. 
+Time domain noise leads to irregular alignments, obfuscates relationships and attenuates inferred coefficients. 
+To handle these challenges, we introduce a regression model that considers stochastic time delays as 'errors' in the time domain.
+For a certain subset of problems, by modelling both prediction and time errors it is possible to outperform traditional models.
+Through a simulated experiment, we demonstrate results that significantly improve upon Ordinary Least Squares regression.
+
 \end{abstract}

### 2021-03-21 documentation/article/article.tex
Abstract refinements.

--- 2021-03-21 documentation/article/article.tex
@@
 \begin{abstract}
-... 
+Systems with stochastic time delay between the input and output present a number of unique challenges. 
+Time domain noise leads to irregular alignments, obfuscates relationships and attenuates inferred coefficients. 
+To handle these challenges, we introduce a regression model that regards stochastic time delay as an 'error' in the time domain.
+For a certain subset of problems, by modelling both prediction \emph{and time} time errors it is possible to outperform traditional models.
+Through a simulated experiment of a univariate problem, we demonstrate results that significantly improve upon Ordinary Least Squares (OLS) regression.
 \end{abstract}

### 2021-03-21 documentation/article/article.tex
Added 'Related Work' section.

@@
 Time domain noise leads to irregular alignments, obfuscates relationships and attenuates inferred coefficients. 
-To handle these challenges, we introduce a regression model that regards stochastic time delay as an 'error' in the time domain.
+To handle these challenges, we introduce a maximum likelihood regression model, that regards stochastic time delay as an 'error' in the time domain.
 For a certain subset of problems, by modelling both prediction \emph{and time} time errors it is possible to outperform traditional models.
@@
 Next, we find the maximum likelihood estimate, for a given set of parameters, considering the time error ($t$-axis) and the regression error ($y$-axis) simultaneously.
+
+\section{Related Work}
+There is extensive literature on time series problems with time delay dependencies. In the statistical literature there a number of treatments which focus on 
+fixed time delay dependencies. For example, Granger Causality is used to determine whether one time series is useful for forecasting another, across some fixed delay.
+Distributed Lag and Dynamic Regression models are able to handle linear and non-linear cause/effect relations that exist across multiple time lags.
+Nonetheless, these models assume a fixed time lag dependency.
+
+In the machine learning literature, there are a number of models which can handle complex dependencies across time. Sequence models such as the RNN, LSTM and GRU 
+are able to generate predictions which incorporate time delays between input and the output variables.
+More recently, attention based models such as the Transformer have become the state-of-the-art for a
+variety of tasks.
+Yet, all of these models have a tacit assumption of a fixed time lag dependency.
+There is also Dynamic Time Warping (DTW) which tries to find the optimal alignment between time series sequences by minimizing the
+distance between the respective inputs. DTW is able to model varying time lag dependencies but is not purposed for regression, instead being mainly used for pattern matching and sequence alignment.
+
+The field of system identification has a variety of literature on dynamic non-linear systems. For example, the NARMAX model can be used to identify 
+non-linear systems with fixed time delays between input and output. At the same time, there are also treatments on systems with disturbances in the input, known as EIV systems. To our knowledge,
+these works do not deal with uncertainty in the time domain.
+
+Finally, recent literature such as Dynamic Time Lag Regresion and Variable Lag Granger causality deal with non-stationary time lag dependencies.
+In other words, the lag structure is assumed to evolve over time.
+
+There has been relatively little work on \emph{stochastic time delays} i.e. time delays which vary randomly. In this paper, we deal with this specific case.
 

### 2021-03-21 documentation/article/article.tex
Refine related work section.

@@
 \section{Related Work}
-There is extensive literature on time series problems with time delay dependencies. In the statistical literature there a number of treatments which focus on 
+There is extensive existing literature on time series problems with time delay dependencies. In the statistical disciplines there a number of treatments which focus on 
 fixed time delay dependencies. For example, Granger Causality is used to determine whether one time series is useful for forecasting another, across some fixed delay.
-Distributed Lag and Dynamic Regression models are able to handle linear and non-linear cause/effect relations that exist across multiple time lags.
+Distributed Lag and Dynamic Regression models are able to handle linear and non-linear cause and effect relations that exist across multiple time lags.
 Nonetheless, these models assume a fixed time lag dependency.
@@
 
-The field of system identification has a variety of literature on dynamic non-linear systems. For example, the NARMAX model can be used to identify 
+The field of system identification also has considerable literature on dynamic non-linear systems. For example, the NARMAX model can be used to identify 
 non-linear systems with fixed time delays between input and output. At the same time, there are also treatments on systems with disturbances in the input, known as EIV systems. To our knowledge,
@@
 
-Finally, recent literature such as Dynamic Time Lag Regresion and Variable Lag Granger causality deal with non-stationary time lag dependencies.
+Finally, recent papers such as Dynamic Time Lag Regression and Variable-Lag Granger Causality deal with non-stationary time lag dependencies.
 In other words, the lag structure is assumed to evolve over time.
 
-There has been relatively little work on \emph{stochastic time delays} i.e. time delays which vary randomly. In this paper, we deal with this specific case.
+There has been little work on \emph{stochastic time delays} i.e. time delays which vary randomly. In this paper, we deal with this specific case.
 

### 2021-03-21 documentation/article/article.tex
Add references v1

@@
 \usepackage{tabularx}
-\bibliography{references}
 
@@
 When extended to time series, a number of time specific complexities arise.
-For the specific case where the input $x$ affects the output $y$, with a random time delay in between, the estimated coefficients (or weights, predictions etc) are significantly attenuated.
+For the specific case where the input $x$ affects the output $y$, with a random time delay in between, the estimated coefficients (or weights, predictions etc) are significantly attenuated \cite{spearman}.
 This attenuation occurs both in traditional statistical models and machine learning models, and for certain problems can be a serious limitation.
@@
 Therefore, these levels must be controlled by insulin injections and a special diet.
-The challenge for many people is that the relationship between the input (insulin) and output (blood sugar) is extremely complex.
+The challenge for many people is that the relationship between the input (insulin) and output (blood sugar) is extremely complex \cite{blood_glucose}.
 The effect of the insulin injection might be observed after $15$, $20$  or $t$ minutes depending on a number of factors, many of which are unknown.
@@
 There is extensive existing literature on time series problems with time delay dependencies. In the statistical disciplines there a number of treatments which focus on 
-fixed time delay dependencies. For example, Granger Causality is used to determine whether one time series is useful for forecasting another, across some fixed delay.
-Distributed Lag and Dynamic Regression models are able to handle linear and non-linear cause and effect relations that exist across multiple time lags.
+fixed time delay dependencies. For example, Granger Causality \cite{granger} is used to determine whether one time series is useful for forecasting another, across some fixed delay.
+Distributed Lag \cite{almon} and Dynamic Regression models (e.g \cite{dynr}) are able to handle linear and non-linear cause and effect relations that occur across multiple time lags.
 Nonetheless, these models assume a fixed time lag dependency.
 
-In the machine learning literature, there are a number of models which can handle complex dependencies across time. Sequence models such as the RNN, LSTM and GRU 
+In the machine learning literature, there are a number of models which can handle complex dependencies across time. Sequence models such as the RNN, LSTM and GRU \cite{ml}
 are able to generate predictions which incorporate time delays between input and the output variables.
 More recently, attention based models such as the Transformer have become the state-of-the-art for a
-variety of tasks.
+variety of tasks, including time series forecasting \cite{deep_transformers}.
 Yet, all of these models have a tacit assumption of a fixed time lag dependency.
-There is also Dynamic Time Warping (DTW) which tries to find the optimal alignment between time series sequences by minimizing the
+There is also Dynamic Time Warping (DTW) \cite{dtw} which tries to find the optimal alignment between time series sequences by minimizing the
 distance between the respective inputs. DTW is able to model varying time lag dependencies but is not purposed for regression, instead being mainly used for pattern matching and sequence alignment.
 
-The field of system identification also has considerable literature on dynamic non-linear systems. For example, the NARMAX model can be used to identify 
-non-linear systems with fixed time delays between input and output. At the same time, there are also treatments on systems with disturbances in the input, known as EIV systems. To our knowledge,
+The field of system identification also has considerable literature on dynamic non-linear systems. For example, the NARX \cite{ml} model can be used to identify 
+non-linear systems with fixed time delays between input and output.
+At the same time, there are also treatments on systems with disturbances in the input \cite{eiv_sysid}, known as EIV systems.
+To our knowledge,
 these works do not deal with uncertainty in the time domain.
 
-Finally, recent papers such as Dynamic Time Lag Regression and Variable-Lag Granger Causality deal with non-stationary time lag dependencies.
+Finally, recent papers such as Dynamic Time Lag Regression \cite{dtlr} and Variable-Lag Granger Causality \cite{Amornbunchornvej_2019} deal with non-stationary time lag dependencies.
 In other words, the lag structure is assumed to evolve over time.
@@
 Ordinary regression analyses (and machine learning models) define the loss function with respect to errors in the $y$ axis only.
-For EIV, errors are considered in both the $y$-axis and the $x$-axis (REF).
-This is useful when there are measurement errors in the independent variable eg.
-because your physical measurements have some degree of random error.
+For EIV, errors are considered in both the $y$-axis and the $x$-axis \cite{deming}.
+This is useful when there are measurement errors in the independent variable e.g.
+because the physical measurements have some degree of random error.
 Similarly, for this problem we assume that we have errors in the $y$-axis and the $t$-axis.
@@
 For the outer parameter optimisation, typical methods such as gradient descent, genetic algorithms and annealing can be used.
-In our implementation, we have used the differential evolution algorithm (scipiy.optimize (REF)) because of its ability to handle noisy objective functions.
+In our implementation, we have used the differential evolution algorithm \cite{diff_ev} (scipiy.optimize \cite{scipy}) because of its ability to handle noisy objective functions \cite{diff_ev}.
 In order to improve convergence, we also standardize all input variables to the range (0,1).
@@
 Next, the model could be extended to include distributed lag structures.
-A distributed lag structure is where past values of the impulse influence future values of the output.
+A distributed lag structure is where past values of the impulse influence future values of the output \cite{almon}.
 Finally, in relating $x(t)$ to $f(x(t))$, the function $f$ could be extended to the modelling of non-linear forms.
@@
 In addition, by incorporating these elements we could begin to explore alternative forms of time series prediction.
-Essentially, existing prediction mechanisms predict the average across time shifts.However, in some situations the predictions could be considered nonsensical.
+Essentially, existing prediction mechanisms predict the average across time shifts.
+However, in some situations the predictions could be considered nonsensical.
 
@@
 
-%\printbibliography
+\bibliography{references}
+\bibliographystyle{ieeetr}
 

### 2021-03-22 documentation/article/article.tex
Refine abstract
Updated example in the article for poisson time shifts. Shifts are now positive only. Added true/estimated shifts.

@@
 Time domain noise leads to irregular alignments, obfuscates relationships and attenuates inferred coefficients. 
-To handle these challenges, we introduce a maximum likelihood regression model, that regards stochastic time delay as an 'error' in the time domain.
-For a certain subset of problems, by modelling both prediction \emph{and time} time errors it is possible to outperform traditional models.
+To handle these challenges, we introduce a maximum likelihood regression model that regards stochastic time delay as an 'error' in the time domain.
+For a certain subset of problems, by modelling both prediction \emph{and} time errors it is possible to outperform traditional models.
 Through a simulated experiment of a univariate problem, we demonstrate results that significantly improve upon Ordinary Least Squares (OLS) regression.
@@
 Firstly, we define a parametric form and some initial parameters to be estimated for the function $f$.
-For simplicity, let's take the univariate linear model where the $f(x(t))$ is parameterised by $\beta$, $\sigma_{\varepsilon}$(error standard deviation)  and $\sigma_{\tau}$ (time error standard deviation).
+For simplicity, let's take the univariate linear model where the $f(x(t))$ is parameterised by $\beta$, $\sigma_{\varepsilon}$(error standard deviation)  and $\lambda_{\tau}$ (time error parameter).
 First, we initialise some starting values for each of these parameters.
@@
 \begin{enumerate}
-    \item [(1)] Initialise the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$ ($\theta_{f}$) for the function $f$
+    \item [(1)] Initialise the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$ ($\theta_{f}$) for the function $f$
     \item [(2)] Find the $\mathcal{T}$ which maximises the likelihood for the given set of parameters:
@@
         \end{enumerate} 
-    \item [(3)] Optimise the parameters $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$ ($\theta_{f}$).
+    \item [(3)] Optimise the parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$($\theta_{f}$).
 
@@
 \vspace{5mm} %5mm vertical space
-The procedure consists of an inner optimisation to find the set of $\mathcal{T}$ and an outer optimisation to find the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\sigma_{\tau}$. 
+The procedure consists of an inner optimisation to find the set of $\mathcal{T}$ and an outer optimisation to find the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$. 
 By iteratively alternating between the time shift optimisation and the parameter optimisation we are able to maximise equation (4). 
@@
 The accuracy of the final parameter estimate is relative to the ratio of the $y$-axis error and the effect size $f(x(t))$.
-As $\sigma_{\varepsilon}/f(x(t))$ $\rightarrow$ $\infty$ , $\sigma_{\tau}$ $\rightarrow$ $0$ (\textcolor{red}{we need to write this as $\lim$ to be more formal}).
-In the event that $\sigma_{\tau}=0$, the model becomes a ‘fixed lag’ model, and for the discrete gaussian case the parameter estimate $\beta$ tends to the standard linear regression coefficient.
+As $\sigma_{\varepsilon}/f(x(t))$ $\rightarrow$ $\infty$ , $\lambda_{\tau}$ $\rightarrow$ $0$ (\textcolor{red}{we need to write this as $\lim$ to be more formal}).
+In the event that $\lambda_{\tau}=0$, the model becomes a ‘fixed lag’ model, and for the discrete gaussian case the parameter estimate $\beta$ tends to the standard linear regression coefficient.
 On the other hand as $\sigma_{\varepsilon}/f(x(t))$ tends to zero, $\beta x(t)$ $\rightarrow f(x(t))$ (\textcolor{red}{we need to write this as $\lim$ to be more formal}).
@@
 
-The input signal has 17 non-zero impulses which have been drawn from the standard normal distribution. The system is 'off' when $x(t) =0$. In other words, when $x(t) = 0$, $f(x(t)) = 0$.
+The input signal has 20 non-zero impulses which have been drawn from the standard normal distribution. The system is 'off' when $x(t) =0$. In other words, when $x(t) = 0$, $f(x(t)) = 0$.
 
-The true shift distribution $\mathcal{T}$ is given by $\mathcal{T} \sim N(0, 2)$. There is one value of $\tau$ for each of the 17 impulses.
+The true shift distribution $\mathcal{T}$ is given by $\mathcal{T} \sim Poisson(\lambda = 2)$. There is one value of $\tau$ for each of the 20 impulses.
 
@@
 \end{center}
-
+\newpage
 The true set of $\mathcal{T}$ is:
@@
 \left(
-\begin{array}{ccc}
--1 & 1 & 0 
+\begin{array}{cccccccccccccccccccc}
+2 & 1 & 2 & 0 & 2 & 3 & 1 & 0 & 2 & 4 & 1 & 2 & 3 & 2 & 1 & 0  & 0 & 1 & 1 & 0 
 \end{array}  
@@
 \end{align}
+\newline
 
@@
 \left(
-\begin{array}{ccc}
--1 & 1 & 0 
+\begin{array}{cccccccccccccccccccc}   
+2 & 1 & 2 & 0 & 2 & 3 & 1 & 1 & 2 & 4 & 1 & 2 & 3 & 2 & 1 & 0  & 0 & 1 & 1 & 0 
 \end{array}  
@@
 \hline
-$\beta$ & 1.96 & 0.26 & 2  \\
+$\beta$ & 2.09 & 0.5 & 2  \\
 \hline
-$Intercept$ & 6.48 & 6.62 & 6.5 \\
+$Intercept$ & 6.53 & 6.62 & 6.5 \\
 \hline
-$\sigma_t$ & 1.51 & $N/A$ & 1.47 \\
+$\lambda_{\tau}$ & 1.54 & $N/A$ & 1.4 \\
 \hline
-$\sigma_e$ & 0.26 & 1.03 & 0.2 \\
+$\sigma_{\epsilon}$ & 0.20 & 1.03 & 0.2 \\
 \end{tabular}
@@
 
-As shown in Table \ref{results_table}, the estimated values for $\beta$ and $\sigma_e$ are significantly improved by taking into account the stochastic time delay noise.
-The estimated $\beta$ using OLS regression is $0.25$, compared to the true value $2$.
+As shown in Table \ref{results_table}, the estimated values for $\beta$ and $\sigma_{\epsilon}$ are significantly improved by taking into account the stochastic time delay noise.
+The estimated $\beta$ using OLS regression is $0.5$, compared to the true value $2$.
 Even a small amount of noise in the time axis causes attenuation,
 limiting the usefulness of OLS for these problems.
-In comparison, the algorithm described herein estimates a $\beta$ of $1.96$, much closer to the true value.
+In comparison, the algorithm described herein estimates $\beta$ to be $2.09$, much closer to the true value.
 

### 2021-03-22 documentation/article/article.tex
Changed wording
Update wording

@@
 
-There has been little work on \emph{stochastic time delays} i.e. time delays which vary randomly. In this paper, we deal with this specific case.
+In this paper, we deal with the specific case of \emph{stochastic time delays} i.e. time delays which vary randomly. There appears to be little existing research on this exact topic.
 
@@
 We want to determine the, functional relationship $f:\mathbb{R} \longrightarrow \mathbb{R}$ to some other (target) variable $y$.
-We further impose the extra condition for $f$ to be injective.
+Furthermore, we impose the condition that $f$ be injective.
 The series $x(t)$ must be stationary with a known value for the point at which $f(x(t)) = 0$.

### 2021-03-22 documentation/article/article.tex
Fix some typos and wording

@@
 
-In this paper, we deal with the specific case of \emph{stochastic time delays} i.e. time delays which vary randomly. There appears to be little existing research on this exact topic.
+In this paper, we deal with the specific case of \emph{stochastic time delays} i.e. time delays which vary randomly. There appears to be little existing research on this topic.
 
@@
 That is, we know at which point the input series has no effect on the output (when it’s off).
-The size of the support of $f(x(t))$ should be small relative to the total vector length (we we will explore this constraint in more detail in the 'Limitations' section).
+The size of the support of $f(x(t))$ should be small relative to the total vector length (we we will explore this condition further in the 'Limitations' section).
 We utilise the terminology {\em impulse} for an individual element of the support of $f$.
@@
 
-Given that each non-zero impulse (row) of $X(t)$ is affected a random time delay (denoted $\tau$), we then model:
+Given that each non-zero impulse (row) of $X(t)$ is affected by a random time delay (denoted $\tau$), we then model:
 
@@
 
-This is particularly problematic if there are multiple impulses within a short period of time, or impulses have a distributed effect over multiple time points.
+This is particularly problematic if there are multiple impulses within a short period of time or impulses have a distributed effect over multiple time points.
 
@@
 \begin{enumerate}
-\item Firstly, smaller shifts are more likely than larger shifts (proportionate to the standard deviation of the $\mathcal{T}$ distribution).
+\item Firstly, smaller shifts are more likely than larger shifts (proportionate to the dispersion of the $\mathcal{T}$ distribution).
 The algorithm should explore the space of smaller shifts more often than larger shifts.
@@
 $f(U_i(t + \tau_i))$ is linearly independent from $f(V_j(t + \tau_j))$, 
-then segments will also be independent and the procedure unique (\textcolor{red}{fix notation}). 
+then the segments will also be independent and the procedure unique (\textcolor{red}{fix notation}). 
 In practice, it may not be possible to guarantee complete independence of every segment, 
@@
 \includegraphics[scale=0.5]{images/tvs_errors.png}
-\caption{Errors...}
+\caption{Fit Residuals}
 \label{fig:tvs_errors}
@@
 \includegraphics[scale=0.6]{images/convergence.png} 
-\caption{Convergence ...}
+\caption{Convergence}
 \label{fig:convergence}
@@
 \hline
-$\beta$ & 2.09 & 0.5 & 2  \\
+$\beta$ & 2.09 & 0.50 & 2.00  \\
 \hline
-$Intercept$ & 6.53 & 6.62 & 6.5 \\
+$Intercept$ & 6.53 & 6.62 & 6.50 \\
 \hline
-$\lambda_{\tau}$ & 1.54 & $N/A$ & 1.4 \\
+$\lambda_{\tau}$ & 1.54 & $N/A$ & 1.40 \\
 \hline
-$\sigma_{\epsilon}$ & 0.20 & 1.03 & 0.2 \\
+$\sigma_{\epsilon}$ & 0.20 & 1.03 & 0.20 \\
 \end{tabular}

### 2021-04-05 documentation/article/article.tex
improve authors section
improve methodology notation

@@
 \title{Modelling Stochastic Time Delay for Regression}
-\author{Aaron Pickering, Dr. Juan Camilo Orduz}
+\author[Aaron]{Aaron Pickering}
+\email{\href{mailto:aaron1rcl@gmail.com}{aaron1rcl@gmail.com}}
+\author[Juan]{Juan Camilo Orduz}
+\email{\href{mailto:juanitorduz@gmail.com}{juanitorduz@gmail.com}}
+\urladdr{\href{https://juanitorduz.github.io/}{juanitorduz.github.io}}
 %\address{}
-\email{aaron1rcl@gmail.com}
-%\urladdr{} 
+
 \date{\today}
@@
 
-% TODO: Rewrite concrete expression of support...is it on x ot t?
--- Aaron: f(x(t)).
-Consider x(t) = 0.1 for t is odd, 1 for t is even.
-If f(x(t)) = 0 for t is odd, 1 for t is even then the series is still valid for our method.
-Let $x: \mathcal{D}_n\subset [a, b] \longrightarrow \mathbb{R}$, where $\mathcal{D}_n$ is a finite set of $n$ points.
-We want to determine the, functional relationship $f:\mathbb{R} \longrightarrow \mathbb{R}$ to some other (target) variable $y$.
-Furthermore, we impose the condition that $f$ be injective.
+\subsection{Model Specification}
+
+Let $x(t)$ be a discrete time series.
+We want to determine the, functional relationship $f:\mathbb{R} \longrightarrow \mathbb{R}$ to some other (target) variable $y$ ob observed data $\{y_i\}_{i=1}^{n}$, that is
+$y = f(x(t))$.
 The series $x(t)$ must be stationary with a known value for the point at which $f(x(t)) = 0$.
 That is, we know at which point the input series has no effect on the output (when it’s off).
-The size of the support of $f(x(t))$ should be small relative to the total vector length (we we will explore this condition further in the 'Limitations' section).
+The size of the support of $f(x(t))$ should be small relative to number of points in the domain (we we will explore this condition further in the 'Limitations' section).
 We utilise the terminology {\em impulse} for an individual element of the support of $f$.
@@
 
-\begin{align}
+\begin{align*}
 x(t) = 
@@
 \right)
-\end{align}
+\end{align*}
 
@@
 
-\begin{align}
+\begin{align*}
 X(t) = 
@@
 
-% TODO fix notation for T
-\begin{align}
+\begin{align}\label{Eqn:ModelSpecification}
 y(t)= f(1^TX(t + \mathcal{T})) + \varepsilon
@@
 
-where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, 
- $X(t)$ is the observed input, 
- $1$ is the vector $(1,1,1...1)$,
- $\varepsilon$ is a noise term and $\mathcal{T}$ is the set of single impulse time delays ($\tau$'s).\\
-
+where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input, $1$ is the vector $(1,1,\cdots, 1)\in 
+\mathbb{R}^{1 \times k}$, where $k$ is the number of impulses and $\varepsilon$ is a noise term and $\mathcal{T}$ is the set of single impulse time delays ($\tau$'s).
+The model definition just asummes that the time shifts are applied to the matrix rows and the we reduce by summing up the columns.
 We assume that $\mathcal{T}$ is not a constant, but rather a random draw from some discrete distribution (e.g.
 discrete gaussian, poisson etc).
-For the discrete gaussian kernel $\mathcal{T}\sim T(\mu, \sigma)$ or for the poisson distribution $\mathcal{T}\sim Pois(\lambda)$ and $\varepsilon \sim N(\mu, \sigma_{\varepsilon})$.
+For the discrete gaussian kernel $\mathcal{T}\sim T(\mu, \sigma)$ or for the poisson distribution $\mathcal{T}\sim Pois(\lambda)$.
+Similarly one can model the errors as  $\varepsilon \sim N(\mu, \sigma_{\varepsilon})$.
 
-In order to find the best estimate of $f$, we would like to find the function $f$ which maximises the joint likelihood of the time-domain shifts and the prediction residuals.
+\subsection{Inference}
+
+In order to find the best estimate of $f$, we would like to find the function $f$ which maximises the joint log-likelihood of the time-domain shifts and the prediction residuals.
 Specifically, we maximise:
@@
 \begin{align}
-    \sum_{i=1}^{n} \log(p(y_i | X_{\tau i} = x_i; \theta_{f}, \tau_i)) + \sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}))
+    \mathcal{L} = 
+    \underbrace{\sum_{i=1}^{n} \log(p(y_i | X_{\tau_i} = x_i; \theta_{f}, \tau_i))}_{\mathcal{L}_1} 
+    + 
+    \underbrace{\sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}))}_{\mathcal{L}_2}
 \end{align}
@@
 where $\theta_{f}$ and $\theta_{\mathcal{T}}$ represent the parameters of the model $f$ and time shifts $\mathcal{T}$ respectively.
+In other words, the $\mathcal{L}_2$ term represents likelihood of time shifts and the $\mathcal{L}_1$ term represents the likelihood of the prediction residuals.
+We maximise these terms simultaneously.
 For simplicity, we assume that the time shift distribution and error distributions are independent.
@@
 
-Before optimisation, the values of each individual time shift ($\tau$) are not known.
+Before optimisation, the values of each individual time shift $\tau$ are not known.
 In addition, the prediction errors $\varepsilon$ can only be calculated if each value of $\tau$ is available (because for each time shift there is a different prediction and hence prediction error).
 Therefore, our algorithm finds the optimum set of time shifts in an inner optimisation, while iteratively searching for the optimum parameters of $f$ in an outer optimisation loop.
-Firstly, we define a parametric form and some initial parameters to be estimated for the function $f$.
-For simplicity, let's take the univariate linear model where the $f(x(t))$ is parameterised by $\beta$, $\sigma_{\varepsilon}$(error standard deviation)  and $\lambda_{\tau}$ (time error parameter).
+Firstly, we define a parametric form $f(x;\theta_f)$ and some initial parameters to be estimated for the function $f$.
+For simplicity, let's take the univariate linear model where the $f(x(t))$ is parameterised by $\beta$, $\sigma_{\varepsilon}$(error standard deviation).
+In addition, we choose a poisson distribution for $\mathcal{T}$, where $\theta_{\mathcal{T}}$ is a parameterised by its mean $\lambda_{\mathcal{T}}$.
+
 First, we initialise some starting values for each of these parameters.

### 2021-04-10 documentation/article/article.tex
Further notation improvements, cleanup.

@@
 
-\title{Modelling Stochastic Time Delay for Regression}
+\title{Modelling Stochastic Time Delay for Regression Analysis}
 \author[Aaron]{Aaron Pickering}
@@
 Let $x(t)$ be a discrete time series.
-We want to determine the, functional relationship $f:\mathbb{R} \longrightarrow \mathbb{R}$ to some other (target) variable $y$ ob observed data $\{y_i\}_{i=1}^{n}$, that is
+We want to determine the functional relationship $f:\mathbb{R} \longrightarrow \mathbb{R}$ to some other (target) variable $y$ on observed data $\{y_i\}_{i=1}^{n}$, that is
 $y = f(x(t))$.
@@
 That is, we know at which point the input series has no effect on the output (when it’s off).
-The size of the support of $f(x(t))$ should be small relative to number of points in the domain (we we will explore this condition further in the 'Limitations' section).
+The size of the support of $f(x(t))$ should be small relative to number of points in the domain (we we will explore this condition further in 'Limitations').
 We utilise the terminology {\em impulse} for an individual element of the support of $f$.
@@
 
-where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input, $1$ is the vector $(1,1,\cdots, 1)\in 
-\mathbb{R}^{1 \times k}$, where $k$ is the number of impulses and $\varepsilon$ is a noise term and $\mathcal{T}$ is the set of single impulse time delays ($\tau$'s).
-The model definition just asummes that the time shifts are applied to the matrix rows and the we reduce by summing up the columns.
-We assume that $\mathcal{T}$ is not a constant, but rather a random draw from some discrete distribution (e.g.
+where $X(t + \mathcal{T})$ is the matrix of time shifted impulses, $X(t)$ is the observed input, $\mathcal{T}$ is the set of single impulse time delays ($\tau$'s), $1$ is the vector $(1,1,\cdots, 1)\in
+\mathbb{R}^{1 \times k}$, where $k$ is the number of impulses and $\varepsilon$ is a noise term.
+
+The model definition represents the application of time shifts to the matrix rows and a subsequent reduction by summation over the columns.
+
+We also assume that $\mathcal{T}$ is not a constant, but rather a random draw from some discrete distribution (e.g.
 discrete gaussian, poisson etc).
@@
 Similarly one can model the errors as  $\varepsilon \sim N(\mu, \sigma_{\varepsilon})$.
-
 \subsection{Inference}
@@
     + 
-    \underbrace{\sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}))}_{\mathcal{L}_2}
+    \underbrace{\sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}, \theta_{f}))}_{\mathcal{L}_2}
 \end{align}
@@
 In addition, the prediction errors $\varepsilon$ can only be calculated if each value of $\tau$ is available (because for each time shift there is a different prediction and hence prediction error).
-Therefore, our algorithm finds the optimum set of time shifts in an inner optimisation, while iteratively searching for the optimum parameters of $f$ in an outer optimisation loop.
+Therefore, our algorithm finds the optimum set of time shifts in an inner optimisation ($\mathcal{L}_2$), while iteratively searching for the optimum parameters of $f$ in an outer optimisation loop ($\mathcal{L}_1$).
 Firstly, we define a parametric form $f(x;\theta_f)$ and some initial parameters to be estimated for the function $f$.
-For simplicity, let's take the univariate linear model where the $f(x(t))$ is parameterised by $\beta$, $\sigma_{\varepsilon}$(error standard deviation).
+For simplicity, let's take the univariate linear model with gaussian errors, where the $f(x(t))$ is parameterised by $\beta$ and $\sigma_{\varepsilon}$ (error standard deviation).
 In addition, we choose a poisson distribution for $\mathcal{T}$, where $\theta_{\mathcal{T}}$ is a parameterised by its mean $\lambda_{\mathcal{T}}$.
+The choice of a poisson distribution for $\mathcal{T}$
+ensures discrete, positive time shifts only.
 
-First, we initialise some starting values for each of these parameters.
+First, let us initialise some starting values for each of these parameters.
 
 Now, we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$.
-It stands to reason that the best possible time shift would be one that is not too far away from the observed impulse (i.e.
-a high likelihood given some distribution) and also produces the best possible prediction.
+It stands to reason that the best possible time shift would be one that is not too distant from the observed impulse (i.e.
+has a high likelihood given some distribution) and also produces the best possible prediction.
 In this example, we can calculate the prediction $y$ by simply multiplying the shifted value by its parameter $\beta$.
@@
 
-\begin{align}
+\begin{align*}
 X(t + \mathcal{T}) = 
@@
 \right)
-\end{align}
+\end{align*}
 
 and the effect is therefore 
-\begin{align}
+\begin{align*}
 y = 
@@
 \begin{enumerate}
-    \item [(1)] Initialise the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$ ($\theta_{f}$) for the function $f$
+    \item [(1)] Initialise the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$ for the function $f$.
     \item [(2)] Find the $\mathcal{T}$ which maximises the likelihood for the given set of parameters:
         \begin{enumerate} 
-            \item[(i)] Initialize the optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution (e.g. zero for the zero mean gaussian). 
+            \item[(i)] Initialize the discrete optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution ($\lambda_{\tau}$).
             \item[(ii)] Next, randomly select a small number of impulses, with the value treated as a hyperparameter.
-The time shifts are then randomly changed to a value drawn for the $\mathcal{T}$ distribution (proposal vector), and the likelihood (both time shift and prediction error) for the proposal is calculated.
-            \item[(iii)] If the proposal likelihood is higher than the current maximum likelihood, we update our estimate.
-            \item [(iv)] Return to (i) and repeat, with the best estimate of the set of $\mathcal{T}$ improving over a number of iterations.
-The number of iterations ($n$) is also treated as a hyperparameter.
+For each impulse, a random time shift is drawn from the $\mathcal{T}$ distribution creating a proposal vector.
+The likelihood (both time shift and prediction error) for the proposal is calculated.
+            \item[(iii)] If the proposal likelihood is higher than the current maximum likelihood, our estimate is updated.
+            \item [(iv)] Return to (i) and repeat. The best estimate of the set of $\mathcal{T}$ improves each iteration. The number of iterations ($n$) is also treated as a hyperparameter.
         \end{enumerate} 
-    \item [(3)] Optimise the parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$($\theta_{f}$).
+    \item [(3)] Optimise the model parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$.
 
@@
 \vspace{5mm} %5mm vertical space
-The procedure consists of an inner optimisation to find the set of $\mathcal{T}$ and an outer optimisation to find the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$. 
-By iteratively alternating between the time shift optimisation and the parameter optimisation we are able to maximise equation (4). 
-For the outer parameter optimisation, typical methods such as gradient descent, genetic algorithms and annealing can be used.
+The procedure consists of an inner optimisation $\mathcal{L}_2$ to find the set of $\mathcal{T}$ and an outer optimisation $\mathcal{L}_1$ to find the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$.
+By iteratively alternating between the time shift optimisation and the parameter optimisation we are able to maximise equation (2). 
+For the outer parameter optimisation, typical methods such as gradient descent, genetic algorithms and simulated annealing can be used.
 In our implementation, we have used the differential evolution algorithm \cite{diff_ev} (scipiy.optimize \cite{scipy}) because of its ability to handle noisy objective functions \cite{diff_ev}.
@@
 
-The accuracy of the final parameter estimate is relative to the ratio of the $y$-axis error and the effect size $f(x(t))$.
-As $\sigma_{\varepsilon}/f(x(t))$ $\rightarrow$ $\infty$ , $\lambda_{\tau}$ $\rightarrow$ $0$ (\textcolor{red}{we need to write this as $\lim$ to be more formal}).
-In the event that $\lambda_{\tau}=0$, the model becomes a ‘fixed lag’ model, and for the discrete gaussian case the parameter estimate $\beta$ tends to the standard linear regression coefficient.
-On the other hand as $\sigma_{\varepsilon}/f(x(t))$ tends to zero, $\beta x(t)$ $\rightarrow f(x(t))$ (\textcolor{red}{we need to write this as $\lim$ to be more formal}).
-The method provides no guarantee on recovering the exact time shifts, only that the estimates are equal to or better than their OLS counterparts.
+We also note that the accuracy of the final parameter estimate is relative to the ratio of the $y$-axis error and the effect size $f(x(t))$. 
+As the ratio of noise $\sigma_{\epsilon}$ to effect $f(x(t))$ increases, the time shift distribution $\mathcal{T}$ shrinks. 
+In the event that $\lambda_{\tau}$ becomes 0, the model becomes a ‘fixed lag’ model, and the parameter estimate $\beta$ tends to the standard linear regression coefficient.
+Therefore, the method provides no guarantee on recovering the exact time shifts, only that the coefficient estimates are equal to or better than their OLS counterparts.
 
 \section{Limitations}
+\subsection{Scaling}
 As the length of $x(t)$ increases, more impulses are introduced and the size of the decomposed matrix $X(t)$ also increases.
@@
 in parallel, making the inner optimisation procedure tractable for longer length sequences. 
+\newline
+\subsection{Constraints}
+Finally, a note on problem constraints.
+The likelihood estimates are dependent on the inner optimisation procedure.
+There is no gaurantee that the global maximum will be found, particularly for sequences with a high density of impulses.
 
-Finally, a note on problem constraints.
-The likelihood estimates are dependent on the inner ($\mathcal{T}$) optimisation procedure.
-This procedure is $NP$-$hard$ and there is no gaurantee that the global maximum will be found, particularly for sequences with a high density of impulses.
+For example, consider the example sequence:
 
-For example, consider the example sequence
-
-\begin{align}
+\begin{align*}
     x(t) = 
@@
 
-The following section demonstrates a simulated univariate example. Full code to reproduce the example can be found via \href{https://github.com/aaron1rcl/tvs_regression/}{TVS Regression on GitHub}.
+For the model, we use the nomenclature Time Varying Stochastic (TVS) Regression and the full code to reproduce the example can be found via \href{https://github.com/aaron1rcl/tvs_regression/}{TVS Regression on GitHub}.
 
-Figure \ref{fig:simulated_example} shows the simulated time series.
+The following section demonstrates a simulated univariate example of TVS Regression. Figure \ref{fig:simulated_example} shows the simulated time series.
 
@@
 
-The green line represents the shifted series, corresponding to the time at which the effect occurs. The blue line is the output $y$.
+The green line represents the shifted series, corresponding to the time at which the effect occurs. The blue line is the output $y$ which includes a small amount of gaussian noise.
 
-The values for the parameters were arbitrarily selected. The output y is defined by the following equation:
+The values for the parameters were arbitrarily selected. The output y is therefore defined by the following equation:
 
@@
 After fitting the model, we obtain the following results.
-Figure \ref{fig:tvs_ols_fit} show the model fit (denoted TVS) and a comparison with standard OLS.
-Figure \ref{fig:tvs_errors} and Figure \ref{fig:convergence} show the errors distribution of the TVS fit and  convergence of the TVS model respectively.
+Figure \ref{fig:tvs_ols_fit} shows the model fit (denoted TVS) and a comparison with standard OLS.
+Figures \ref{fig:tvs_errors} and Figure \ref{fig:convergence} show the error distribution of the TVS fit and the convergence of the TVS model respectively.
 
@@
 
-%To-do - write in the actual T distributions above. Not sure how to fix the hbox problem.
-
-%To-do properly format the table
 \begin{center}
@@
 limiting the usefulness of OLS for these problems.
-In comparison, the algorithm described herein estimates $\beta$ to be $2.09$, much closer to the true value.
+In comparison, the TVS Regression algorithm estimates $\beta$ to be $2.09$, much closer to the true value.
 
@@
 
-%To-DO Further questions:
- %- what to write in the abstract?
- %- we havent addressed previous research that covers related problems, shall we include a section?
- %- add references for all unbacked major statements
- %- should we address identifiability?
- %- examplte notebook should be fully aligned with the paper - fix
-
 \end{document}

### 2021-04-11 documentation/article/article.tex
Edited future work section.
improve notation Algorithm section

@@
 \usepackage{tabularx}
+
+\theoremstyle{definition}
+\newtheorem*{example}{Example}
+\theoremstyle{definition}
+\newtheorem*{remark}{Remark}
 
@@
     + 
-    \underbrace{\sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}, \theta_{f}))}_{\mathcal{L}_2}
+    \underbrace{\sum_{i=1}^{n} \log(p(\tau_i| \theta_{\mathcal{T}}))}_{\mathcal{L}_2}
 \end{align}
@@
 Firstly, we define a parametric form $f(x;\theta_f)$ and some initial parameters to be estimated for the function $f$.
-For simplicity, let's take the univariate linear model with gaussian errors, where the $f(x(t))$ is parameterised by $\beta$ and $\sigma_{\varepsilon}$ (error standard deviation).
+
+\begin{example}\label{example:poisson}
+Let's take the univariate linear model with gaussian errors, where the $f(x(t))$ is parameterised by $\beta$ and $\sigma_{\varepsilon}$ (error standard deviation).
 In addition, we choose a poisson distribution for $\mathcal{T}$, where $\theta_{\mathcal{T}}$ is a parameterised by its mean $\lambda_{\mathcal{T}}$.
@@
 ensures discrete, positive time shifts only.
+\end{example}
 
-First, let us initialise some starting values for each of these parameters.
-
+First, let us iniialise some starting values for each of these parameters.
 Now, we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$.
@@
 has a high likelihood given some distribution) and also produces the best possible prediction.
-In this example, we can calculate the prediction $y$ by simply multiplying the shifted value by its parameter $\beta$.
 From there, the likelihood estimate is derived from the time shift and the prediction error.
-
 Finally, we iterate over a number of values of $\tau$ optimising until we maximise the likelihood for the specific impulse.
 
-However, we must also consider that the impulses in $X(t)$ are not independent from each other. After shifting, it's possible that two or more effects occur simultaneously.
-
+\begin{remark}
+However, we must also consider that the impulses in $X(t)$ are not independent from each other.
+After shifting, it's possible that two or more effects occur simultaneously.
 This is particularly problematic if there are multiple impulses within a short period of time or impulses have a distributed effect over multiple time points.
-
 As an example consider the series
@@
 \end{align*}
+\end{remark}
 
@@
 For the optimisation step we utilise two assumptions.
+
 \begin{enumerate}
@@
 
-Accordingly, the optimisation procedure is:
-\vspace{5mm} %5mm vertical space
+Accordingly, the optimisation procedure is (CHANGE TO GENERAL PARAMETERS):
 
@@
             \item[(i)] Initialize the discrete optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution ($\lambda_{\tau}$).
-            \item[(ii)] Next, randomly select a small number of impulses, with the value treated as a hyperparameter.
+            \item[(ii)] Next, randomly select a small number of impulses $m$, with the value of $m$ treated as a hyperparameter.
 For each impulse, a random time shift is drawn from the $\mathcal{T}$ distribution creating a proposal vector.
-The likelihood (both time shift and prediction error) for the proposal is calculated.
-            \item[(iii)] If the proposal likelihood is higher than the current maximum likelihood, our estimate is updated.
-            \item [(iv)] Return to (i) and repeat. The best estimate of the set of $\mathcal{T}$ improves each iteration. The number of iterations ($n$) is also treated as a hyperparameter.
+            \item [(iii)]The likelihood (both time shift and prediction error) for the proposal is calculated. 
+            \item[(iv)] If the proposal likelihood is higher than the current maximum likelihood, our estimate is updated.
+            \item [(v)] Return to (i) and repeat. The best estimate of the set of $\mathcal{T}$ improves each iteration. The number of iterations ($N$) is also treated as a hyperparameter.
         \end{enumerate} 
     \item [(3)] Optimise the model parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$.
+\end{enumerate}
 
-\end{enumerate}
-\vspace{5mm} %5mm vertical space
-The procedure consists of an inner optimisation $\mathcal{L}_2$ to find the set of $\mathcal{T}$ and an outer optimisation $\mathcal{L}_1$ to find the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$.
-By iteratively alternating between the time shift optimisation and the parameter optimisation we are able to maximise equation (2). 
-For the outer parameter optimisation, typical methods such as gradient descent, genetic algorithms and simulated annealing can be used.
+\begin{remark}
+In Example \ref{example:poisson} we recommend to initialize the discrete optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution ($\lambda_{\tau}$).
+\end{remark}
+
+For the outer parameter optimisation (Step (3)), typical methods such as gradient descent, genetic algorithms and simulated annealing can be used.
 In our implementation, we have used the differential evolution algorithm \cite{diff_ev} (scipiy.optimize \cite{scipy}) because of its ability to handle noisy objective functions \cite{diff_ev}.
@@
 
+\begin{remark}
 We also note that the accuracy of the final parameter estimate is relative to the ratio of the $y$-axis error and the effect size $f(x(t))$. 
-As the ratio of noise $\sigma_{\epsilon}$ to effect $f(x(t))$ increases, the time shift distribution $\mathcal{T}$ shrinks. 
-In the event that $\lambda_{\tau}$ becomes 0, the model becomes a ‘fixed lag’ model, and the parameter estimate $\beta$ tends to the standard linear regression coefficient.
+As the ratio of noise $\sigma_{\varepsilon}$ to effect $f(x(t))$ increases, the time shift distribution $\mathcal{T}$ shrinks. 
+In the event that $\lambda_{\tau}$ (CHANGE to $\theta_{f}$) becomes 0, the model becomes a ‘fixed lag’ model, and the parameter estimate $\beta$ tends to the standard linear regression coefficient.
 Therefore, the method provides no guarantee on recovering the exact time shifts, only that the coefficient estimates are equal to or better than their OLS counterparts.
+\end{remark}
 
@@
 To account for this, we assume that distant impulses do not affect each other.
-Concretely, for the $i_{th}$ and $j_{th}$ impulse (row) in $X(t)$, separated by time $t \longrightarrow \infty$, the effect vectors $f(X_i(t + \tau_i))$ and $f(X_j(t + \tau_j))$ are linearly independent.
+Concretely, for the $i_{th}$ and $j_{th}$ impulse (row) in $X(t)$ are far away then effect vectors are orthogonal.
+Hence, we wan decompose wth matrix $X(t)$ in orthogonal blocks on which the algorithm can then be run separately.
 
-Therefore, the matrix $X(t)$ is decomposed into a number of segments, where each segment is 
-composed of a smaller number of dependent impulses.
-To select the right split locations, we search the time series for locations with the lowest impulse density.
-Given that we wish to split $X(t)$ into independent matrices $U_n(t)$ and $V_m(t)$,
-if we choose the split locations such that $\forall_i$ $\epsilon$ $n$ and  $\forall_j$ $\epsilon$ $m$, 
-$f(U_i(t + \tau_i))$ is linearly independent from $f(V_j(t + \tau_j))$, 
-then the segments will also be independent and the procedure unique (\textcolor{red}{fix notation}). 
-In practice, it may not be possible to guarantee complete independence of every segment, 
-in which case results may vary. After segmentation, each matrix can be optimised independently 
-in parallel, making the inner optimisation procedure tractable for longer length sequences. 
-\newline
-\subsection{Constraints}
+\subsection{Constraints (TO DO: ReWrite)}
 Finally, a note on problem constraints.
 The likelihood estimates are dependent on the inner optimisation procedure.
-There is no gaurantee that the global maximum will be found, particularly for sequences with a high density of impulses.
-
-For example, consider the example sequence:
+There is no guarantee that the global maximum will be found, particularly for sequences with a high density of impulses.
+For example, consider the sequence:
 
@@
 
-In the univariate case, the utility of the described method is limited.
-However, we propose three extensions as future work.
-The first is to extend the method to multiple regression.
+In the univariate case, the utility of the described method is limited. However, we propose three extensions as future work. The first is to extend the method to multiple regression. 
 We believe that the extension can be built on the same fundamental ideas presented in this document.
@@
 
-Through these extensions we can begin to tackle a larger number of practical problems.
+Through these extensions we can begin to tackle a number of practical problems that are defined by stochastic relationships between the input and output e.g.
+the blood glucose problem described in the introduction.
 

### 2021-04-11 src/model/linearTVSR.py
Segmentation capability added.

@@ -74,6 +74,7 @@ class linearTVSRModel:
     
             # Scale the time dimension to have a max value of 1
             if max_t is None:
+                # TO-DO: remove hardcoded 5% value and replace with something general
                 #Make the max t 5% of the length by default
                 self.scale_t = self.x.shape[0]*0.05
             else:
@@ -127,8 +128,8 @@ class linearTVSRModel:
         sd = params[2]
         C = params[3]
         
-        # Hack workaround for values outside of the range
-        # Need a cleaner solution
+        # TO-DO: Currently this is a quick hack workaround for values outside of the range
+        # Need a clean general solution.
         if tsd <= 0:
             return 10000
         
@@ -141,10 +142,12 @@ class linearTVSRModel:
         # Apply some rules to the loop boundaries
         loop_bounds = src.refine_bounds(np.copy(self.X), tsd)
 
-        ##### TO-DO
+        #Either split into orthogonal segments or not
         if self.split == True:
             # Define an input to select whether to split or not
-            X_segs, y_segs, dim_segs, bnds_segs = create_segments(np.copy(self.X), np.copy(self.x), np.copy(self.y), loop_bounds)
+            X_segs, y_segs, dim_segs, bnds_segs = create_segments(np.copy(self.X), np.copy(self.x), np.copy(self.y), loop_bounds, tsd)
+            self.segs = dim_segs
+ 
             # Loop through all the segments
             vals = np.array([])
             taus = np.array([])
@@ -163,6 +166,7 @@ class linearTVSRModel:
                 val, tau = self.inner_optimisation(dim_seg, f)
                 
                 print("------ Seg Likelihood: " + str(val))
+                # Add the segment values a list
                 vals = np.append(vals, val)
                 taus = np.append(taus, tau)
         else:
@@ -214,15 +218,16 @@ class linearTVSRModel:
             
     def inner_optimisation(self,  dim, f):
         # Run the inner optimisation algorithm and retrieve the max likelihood
-        val, tau = tau_optimiser([0]*dim[0], f, 1000, 3)
+        val, tau = tau_optimiser([0]*dim[0], f, self.iter, 3)
 
         return val, tau
             
             
-    def fit(self, x_train, y_train, method="L-BFGS-B", split=True, max_t=None):
+    def fit(self, x_train, y_train, method="L-BFGS-B", split=True, max_t=None, iterations=1000):
         
         # Define the setting
         self.split = split
+        self.iter = iterations
         
         # Load the data into the object memory and decompose the vector
         self.x = x_train

### 2021-04-11 src/optimisation/segmentation.py
Added segmentation capability.

@@ -1,10 +1,10 @@
 import src.support as src
 import numpy as np
 
-
-def create_segments(X, x, y, bounds):        
+ 
+def create_segments(X, x, y, bounds, tsd ):        
     loc, loc_diff = get_impulse_locations(x)
-    splits = get_split_points(bounds, loc, loc_diff)
+    splits = get_split_points(bounds, loc, loc_diff, tsd)
     # Split the dependent variable
     y_segments = np.split(y, splits)
     
@@ -35,15 +35,12 @@ def get_impulse_locations(x):
     return loc, loc_diff
     
     
-def get_split_points(bounds, loc, loc_diff):
-    
-    # Extract the maximum boundary
-    global_max_bound = np.max(bounds)
-    global_min_bound = np.min(bounds)
+def get_split_points(bounds, loc, loc_diff, tsd):
     
 
     # Find the spaces which are greater than twice the boundary (2 points movin toward each other)
-    splits = np.where(loc_diff > np.max([np.abs(global_min_bound), global_max_bound])*2)
+    # To-Do: remove hardcoded hyperparameters
+    splits = np.where(loc_diff > np.min([tsd*3,20]))
     # Fit the array location where these splits occur
     split_locations = loc[splits] + np.round(loc_diff[splits]/2)
         

### 2021-04-11 src/model/linearTVSR.py
Segmentation (#21)
* Segmentation capability added.

* Added segmentation capability.

@@ -75,6 +75,7 @@ class linearTVSRModel:
     
             # Scale the time dimension to have a max value of 1
             if max_t is None:
+                # TO-DO: remove hardcoded 5% value and replace with something general
                 #Make the max t 5% of the length by default
                 self.scale_t = self.x.shape[0]*0.05
             else:
@@ -128,8 +129,8 @@ class linearTVSRModel:
         sd = params[2]
         C = params[3]
         
-        # Hack workaround for values outside of the range
-        # Need a cleaner solution
+        # TO-DO: Currently this is a quick hack workaround for values outside of the range
+        # Need a clean general solution.
         if tsd <= 0:
             return 10000
         
@@ -142,10 +143,12 @@ class linearTVSRModel:
         # Apply some rules to the loop boundaries
         loop_bounds = src.refine_bounds(np.copy(self.X), tsd)
 
-        ##### TO-DO
+        #Either split into orthogonal segments or not
         if self.split == True:
             # Define an input to select whether to split or not
-            X_segs, y_segs, dim_segs, bnds_segs = create_segments(np.copy(self.X), np.copy(self.x), np.copy(self.y), loop_bounds)
+            X_segs, y_segs, dim_segs, bnds_segs = create_segments(np.copy(self.X), np.copy(self.x), np.copy(self.y), loop_bounds, tsd)
+            self.segs = dim_segs
+ 
             # Loop through all the segments
             vals = np.array([])
             taus = np.array([])
@@ -164,6 +167,7 @@ class linearTVSRModel:
                 val, tau = self.inner_optimisation(dim_seg, f)
                 
                 print("------ Seg Likelihood: " + str(val))
+                # Add the segment values a list
                 vals = np.append(vals, val)
                 taus = np.append(taus, tau)
         else:
@@ -215,15 +219,16 @@ class linearTVSRModel:
             
     def inner_optimisation(self,  dim, f):
         # Run the inner optimisation algorithm and retrieve the max likelihood
-        val, tau = tau_optimiser([0]*dim[0], f, 1000, 3, self.family)
+        val, tau = tau_optimiser([0]*dim[0], f, self.iter, 3, self.family)
 
         return val, tau
             
             
-    def fit(self, x_train, y_train, method="L-BFGS-B", split=True, max_t=None, family=None):
+    def fit(self, x_train, y_train, method="L-BFGS-B", split=True, max_t=None, family=None, iterations=1000):
         
         # Define the setting
         self.split = split
+        self.iter = iterations
         
         # Set the family
         if family is not None:

### 2021-04-18 documentation/article/article.tex
improve bib, fix table and add correct examples and remarks style
final comments on conclusion and future work

@@
 \theoremstyle{definition}
-\newtheorem*{example}{Example}
-\theoremstyle{definition}
-\newtheorem*{remark}{Remark}
+\newtheorem{example}{Example}[section]
+\newtheorem{remark}{Remark}[section]
 
@@
 \title{Modelling Stochastic Time Delay for Regression Analysis}
-\author[Aaron]{Aaron Pickering}
+\author[Aaron Pickering]{Aaron Pickering}
 \email{\href{mailto:aaron1rcl@gmail.com}{aaron1rcl@gmail.com}}
-\author[Juan]{Juan Camilo Orduz}
+\author[Juan Camilo Orduz]{Juan Camilo Orduz}
 \email{\href{mailto:juanitorduz@gmail.com}{juanitorduz@gmail.com}}
@@
 \date{\today}
-\maketitle
 
@@
 \end{abstract}
+
+\maketitle
 
@@
 We treat the stochastic time delay components as an ‘error’ in the time axis.
-Next, we find the maximum likelihood estimate, for a given set of parameters, considering the time error ($t$-axis) and the regression error ($y$-axis) simultaneously.
+Next, we find the maximum likelihood estimate, for a given set of parameters, considering the time error ($t$-axis) and the regression error ($y$-axis) simultaneously.\\
+
+An open source Python implementation is available on \href{https://github.com/aaron1rcl/tvs_regression/}{GitHub}.
 
@@
 For simplicity, we assume that the time shift distribution and error distributions are independent.
+\\
+
+We will refer to this algorithm as {\em Time Varying Stochastic (TVS) Regression}.
 
@@
 
-For the model, we use the nomenclature Time Varying Stochastic (TVS) Regression and the full code to reproduce the example can be found via \href{https://github.com/aaron1rcl/tvs_regression/}{TVS Regression on GitHub}.
+The following section demonstrates a simulated univariate example of TVS Regression.
+Figure \ref{fig:simulated_example} shows the simulated time series.
+The full code can be found via \href{https://github.com/aaron1rcl/tvs_regression/blob/master/notebooks/1_univariate_example.ipynb}{TVS Regression on GitHub}.
+The input signal has $20$ non-zero impulses which have been drawn from the standard normal distribution.
+The system is 'off' when $x(t) =0$.
+In other words, when $x(t) = 0$, $f(x(t)) = 0$.
+The true shift distribution $\mathcal{T}$ is given by $\mathcal{T} \sim Poisson(\lambda = 2)$.
+There is one value of $\tau$ for each of the 20 impulses.
+The green line in Figure \ref{fig:simulated_example} represents the shifted series, corresponding to the time at which the effect occurs.
+The blue line is the output $y$ which includes a small amount of gaussian noise.
+The values for the parameters were arbitrarily selected.
+The output $y$ is defined by the following equation:
 
-The following section demonstrates a simulated univariate example of TVS Regression. Figure \ref{fig:simulated_example} shows the simulated time series.
+\begin{align*}
+    y = 2(1^TX(t + \mathcal{T})) + 6.5
+\end{align*}
 
-The input signal has 20 non-zero impulses which have been drawn from the standard normal distribution. The system is 'off' when $x(t) =0$. In other words, when $x(t) = 0$, $f(x(t)) = 0$.
-
-The true shift distribution $\mathcal{T}$ is given by $\mathcal{T} \sim Poisson(\lambda = 2)$. There is one value of $\tau$ for each of the 20 impulses.
-
-The green line represents the shifted series, corresponding to the time at which the effect occurs. The blue line is the output $y$ which includes a small amount of gaussian noise.
-
-The values for the parameters were arbitrarily selected. The output y is therefore defined by the following equation:
-
-\begin{align}
-    y = 2(1^TX(t + \mathcal{T})) + 6.5
-\end{align}
+A histogram of the actual distribution of $\mathcal{T}$ is shown in Figure \ref{fig:real_shifts}.
 
@@
 
-A histogram of the actual distribution of $\mathcal{T}$ is shown in Figure \ref{fig:real_shifts}.
+After fitting the model, we obtain the following results.
+Figure \ref{fig:tvs_ols_fit} shows the model fit (denoted TVS) and a comparison with standard OLS.
+Figures \ref{fig:tvs_errors} and Figure \ref{fig:convergence} show the error distribution of the TVS fit and the convergence of the TVS model respectively.
 
@@
 \end{center}
-
-After fitting the model, we obtain the following results.
-Figure \ref{fig:tvs_ols_fit} shows the model fit (denoted TVS) and a comparison with standard OLS.
-Figures \ref{fig:tvs_errors} and Figure \ref{fig:convergence} show the error distribution of the TVS fit and the convergence of the TVS model respectively.
 
@@
 \end{center}
-\newpage
-The true set of $\mathcal{T}$ is:
 
-\begin{align}
-\mathcal{T} = 
+The true values of $\mathcal{T}$ (true shifts) and the estimated values $\mathcal{T}_{est}$ are:
+
+\begin{align*}
+\mathcal{T} &= 
 \left(
@@
 \end{array}  
-\right)
-\end{align}
-\newline
-
-Compared to the estimated $\mathcal{T}$:
-
-\begin{align}
-\mathcal{T}_{est} = 
+\right) \\
+\mathcal{T}_{est} &= 
 \left(
@@
 \begin{table}
-\begin{tabular}{c|c|c|c|}
+\begin{tabular}{|c|c|c|c|}
 \hline
@@
 $\sigma_{\epsilon}$ & 0.20 & 1.03 & 0.20 \\
+\hline
 \end{tabular}
-\caption{ ... } 
+\caption{ Parameter comparison between TVR and OLS regression against the true value.} 
 \label{results_table}
@@
 
-As shown in Table \ref{results_table}, the estimated values for $\beta$ and $\sigma_{\epsilon}$ are significantly improved by taking into account the stochastic time delay noise.
+As shown in Table \ref{results_table}, the estimated values for $\beta$ and $\sigma_{\varepsilon}$ are significantly improved by taking into account the stochastic time delay noise.
 The estimated $\beta$ using OLS regression is $0.5$, compared to the true value $2$.
-Even a small amount of noise in the time axis causes attenuation,
-limiting the usefulness of OLS for these problems.
-In comparison, the TVS Regression algorithm estimates $\beta$ to be $2.09$, much closer to the true value.
+Even a small amount of noise in the time axis causes attenuation limiting the usefulness of OLS for these problems.
+In comparison, the TVS Regression algorithm estimates $\beta$ to be $2.09$, much closer to the true value.\\
 
 While the example is based on simulated data, we believe that the experiment demonstrates clear potential for improvement in the modelling of real world systems with stochastic time delay noise.
-
-\section{Future Work}
-
-In the univariate case, the utility of the described method is limited. However, we propose three extensions as future work. The first is to extend the method to multiple regression. 
-We believe that the extension can be built on the same fundamental ideas presented in this document.
-Next, the model could be extended to include distributed lag structures.
-A distributed lag structure is where past values of the impulse influence future values of the output \cite{almon}.
-Finally, in relating $x(t)$ to $f(x(t))$, the function $f$ could be extended to the modelling of non-linear forms.
-
-Through these extensions we can begin to tackle a number of practical problems that are defined by stochastic relationships between the input and output e.g.
-the blood glucose problem described in the introduction.
-
-In addition, by incorporating these elements we could begin to explore alternative forms of time series prediction.
-Essentially, existing prediction mechanisms predict the average across time shifts.
-However, in some situations the predictions could be considered nonsensical.
-
-For example, let us consider a regression model to predict how much coffee someone drinks between 9am and 9:30am, at work.
-They set off every day in their car at exactly 8:30am and it takes about 30 minutes ($\pm \tau$) to get to work.
-The time shift $\tau$ could be caused by something as simple as variations in traffic.
-When they get to work, they always have a coffee.
-A typical regression might predict $0.5$ coffees between 8:30 am and 9am and $0.5$ coffees between 9am and 9:30am.
-Alternative prediction forms such as: 'When they get to work, they will have 1 coffee' might be more useful.
 
@@
 Our approach allows for consistently improved estimation and prediction when the input is affected by noise in the time domain.
-To our knowledge, the method is novel for this class of problem.
+To our knowledge, the method is novel for this class of problem.\\
 
-%To-DO: Write References
+We propose two extensions as future work.
+The first is to extend the method to multiple regression.
+We believe that the extension can be built on the same fundamental ideas presented in this document.
+Next, the model could be extended to include distributed lag structures.
+A distributed lag structure is where past values of the impulse influence future values of the output \cite{almon}.
+Through these extensions we can begin to tackle a number of practical problems that are defined by stochastic relationships between the input and output e.g.
+the blood glucose problem described in the introduction.
+In addition, we could also begin to apply the model to forecasting problems where the inputs are affected by stochastic time delay.
 
+% References
+\bibliographystyle{acm}
 \bibliography{references}
-\bibliographystyle{ieeetr}
 

### 2021-04-20 documentation/article/article.tex

@@
 
-\title{Modelling Stochastic Time Delay for Regression Analysis}
+\title{Modelling Stochastic Time Delay for Regression}
 \author[Aaron Pickering]{Aaron Pickering}
-\email{\href{mailto:aaron1rcl@gmail.com}{aaron1rcl@gmail.com}}
+\email{\href{mailto:aaron_rcl@hotmail.com}{aaron\_rcl@hotmail.com}}
 \author[Juan Camilo Orduz]{Juan Camilo Orduz}
@@
 \includegraphics[totalheight=6cm]{images/2_input_delay_example.png}
-\caption{Add figure caption}
+\caption{Stochastic Time Delay Example}
 \label{fig:verticalcell}
@@
 
-First, let us iniialise some starting values for each of these parameters.
+First, let us initialise some starting values for each of these parameters.
 Now, we want to find the best possible time shift $\tau$ for each input impulse in $X(t)$.
@@
 
-Accordingly, the optimisation procedure is (CHANGE TO GENERAL PARAMETERS):
+Accordingly, the optimisation procedure is:
 
 \begin{enumerate}
-    \item [(1)] Initialise the set of parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$ for the function $f$.
+    \item [(1)] Initialise the set of parameters $\theta_{f}$ and $\theta_{\mathcal{T}}$ for the function $f$.
     \item [(2)] Find the $\mathcal{T}$ which maximises the likelihood for the given set of parameters:
         \begin{enumerate} 
-            \item[(i)] Initialize the discrete optimisation algorithm with all parameters set to the mean of the $\mathcal{T}$ distribution ($\lambda_{\tau}$).
+            \item[(i)] Initialize the discrete optimisation algorithm time shifts ($\mathcal{T}$). 
             \item[(ii)] Next, randomly select a small number of impulses $m$, with the value of $m$ treated as a hyperparameter.
@@
         \end{enumerate} 
-    \item [(3)] Optimise the model parameters $\beta$, $\sigma_{\varepsilon}$ and $\lambda_{\tau}$.
+    \item [(3)] Optimise the model parameters $\theta_{f}$ and $\theta_{\mathcal{T}}$.
 \end{enumerate}
@@
 
-\subsection{Constraints (TO DO: ReWrite)}
+\subsection{Constraints}
 Finally, a note on problem constraints.
@@
 There is no guarantee that the global maximum will be found, particularly for sequences with a high density of impulses.
-For example, consider the sequence:
-
-\begin{align*}
-    x(t) = 
-    \left(
-\begin{array}{cccccccccc}
-    0.5 & 0.6 & 0.4 & 1 & 2 & 1 &
-    0.3 & 0.2 & 0.5 & 0.8 
-\end{array}  
-\right)
-\end{align*}
-
-where each impulse can shift between -2 and +2 positions. Ignoring shifts beyond the edge of the vector, the solutions space of $\mathcal{T}$ is $5^t$ combinations.  
-In such cases, the estimated likelihood is likely to be close to but not exactly the same as the real value.
-As the density (in time) of impulses increases, the accuracy of the estimates decrease.
-Therefore, this model is most appropriate for sparse time series inputs.
+When the density (in time) of non-zero impulses in $x(t)$ is high, the possible solution space of $\mathcal{T}$ grows exponentially.
+In such cases, the estimated likelihood is likely to be close to, but not exactly the same as the real value. Therefore, TVS Regression is most appropriate for sparse time series inputs.
 
@@
 The green line in Figure \ref{fig:simulated_example} represents the shifted series, corresponding to the time at which the effect occurs.
-The blue line is the output $y$ which includes a small amount of gaussian noise.
+The blue line is the output $y$ which includes a small amount of gaussian noise and an intercept.
 The values for the parameters were arbitrarily selected.
@@
 \end{tabular}
-\caption{ Parameter comparison between TVR and OLS regression against the true value.} 
+\caption{ Parameter comparison between TVS and OLS regression against the true value.} 
 \label{results_table}

### 2021-04-20 documentation/article/article.tex
Add references.bib and plots.

@@
 As the ratio of noise $\sigma_{\varepsilon}$ to effect $f(x(t))$ increases, the time shift distribution $\mathcal{T}$ shrinks. 
-In the event that $\lambda_{\tau}$ (CHANGE to $\theta_{f}$) becomes 0, the model becomes a ‘fixed lag’ model, and the parameter estimate $\beta$ tends to the standard linear regression coefficient.
+In the event that the mean of $\theta_{\mathcal{T}}$ become 0, the model becomes a ‘fixed lag’ model, and the parameter estimate $\beta$ tends to the standard linear regression coefficient.
 Therefore, the method provides no guarantee on recovering the exact time shifts, only that the coefficient estimates are equal to or better than their OLS counterparts.
@@
 To account for this, we assume that distant impulses do not affect each other.
-Concretely, for the $i_{th}$ and $j_{th}$ impulse (row) in $X(t)$ are far away then effect vectors are orthogonal.
-Hence, we wan decompose wth matrix $X(t)$ in orthogonal blocks on which the algorithm can then be run separately.
+Concretely, if the $i_{th}$ and $j_{th}$ impulse (row) in $X(t)$ are far away then their effect vectors are orthogonal.
+Hence, we can decompose the matrix $X(t)$ into orthogonal blocks on which the inner optimisation can be run in parallel.
 
